### 2017-10-01 output/breed_paper.Rnw
Revisions from Bruce, Dave, Chris, and Jim

\documentclass[12pt,]{elsarticle}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{longtable,booktabs}
\usepackage{comment}
\usepackage{graphicx,grffile, textcomp}
\usepackage{setspace}
\usepackage{etoolbox}
\usepackage{float}
\doublespacing
\usepackage{lineno}
\usepackage{lscape} %% Harvard
\bibliographystyle{model2-names}\biboptions{authoryear}
\pagenumbering{gobble}
%% For figure numbers in supplementary info
%% from http://bytesizebio.net/2013/03/11/adding-supplementary-tables-and-figures-in-latex/
\newcommand{\beginsupplement}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
}

\newcommand{\tsups}{\textsuperscript}
\usepackage{Sweave}
\SweaveOpts{prefix.string=graphics/report, png=TRUE}

\begin{document}

\begin{frontmatter}

\title{Rice yield improvements through plant breeding are offset by
  inherent yield declines over time}
\author[ucd]{Matthew B. Espe\corref{corresauthor}}
\cortext[corresauthor]{Corresponding author} \ead{mespe@ucdavis.edu}
\author[ucd]{David Mackill}
\author[ucd]{Jim E. Hill}
\author[ucce2]{Michelle Leinfelder-Miles}
\author[ucce]{Luis A. Espino}
\author[ucce]{Randall Mutters}
\author[ucd]{Chris van Kessel}
\author[ucd]{Bruce A. Linquist}

\address[ucd]{Dept. of Plant Sciences, University of California --
Davis, Davis CA 95616, USA}

\address[ucce]{University of California Cooperative Extension,
Oroville, CA 95965, USA}

\address[ucce2]{University of California Cooperative Extension,
Stockton, CA 95206, USA}

\begin{abstract}

Meeting the challenge of feeding a growing population with
increasingly limited resources will require increasing the yield
potential of staple crops, such as rice. Yet many high-yielding,
intensive production systems have experienced slow rates of yield
improvement in recent years despite a demonstrated increase in the
yield potential of new crop cultivars. We analyzed experimental data
from one such cropping system, i.e., California (CA) rice, in order to
better understand what improvements have been made in the genetic
yield potential obtained through plant breeding. Specifically, the
hypothesis was tested that the yield of a rice cultivar declines over
time after its introduction. This would explain how
side-by-side comparisons show increased yield while no substantive
increases have been seen in farmers' fields. If there is yield
erosion, side-by-side comparisons will show ``false'' yield increases due to the
yield of the older cultivar decreasing, not due to the increased
genetic yield potential of the newer cultivar. An apparent increase in
yield in side-by-side comparisons between old and new rice varieties
was observed when the yield erosion was not considered. However, this
apparent increase in yield could be explained by the differences in
the time when these cultivars were selected. Yield
eroded at an estimated rate of 29.3 kg ha\tsups{-1} year\tsups{-1} (90\%
credible interval -4.4 to -53.3) after initial selection. Once this
effect was considered, the yield advantage of newer cultivars
over old was uncertain (-3.3 kg ha\tsups{-1} year\tsups{-1}, 90\%
credible interval -36.1 to 31.5). Thus, without continuous plant improvement and
deployment of new cultivars, yield will not only remain stagnant but
will decline over time. To increasing the genetic yield potential of
rice in the face of this phenomenon, higher yield targets are needed
to counter-act the decrease over time in the current cultivars.

\end{abstract}

\begin{keyword} rice, yield potential, breeding, yield erosion
\end{keyword}

\end{frontmatter}

\newpage

\textbf{Abbreviations}

CA: California

CI: Credible Interval

RES: Rice Experiment Station

CCRRF: California Cooperative Rice Research Foundation

\newpage

% \linenumbers

<<echo=FALSE>>=

library(rstan)
library(rstanarm)
library(lme4)
library(xtable)

load("../data/model_data.rda")
load("h1_fit.Rda")
load("bivar_fit.Rda")
load("../data/all_vt_weather.Rda")

prep_plot_data = function(x, idx)
{
    as.data.frame(do.call(rbind, tapply(x, idx, function(x){
        c(mean = mean(x), se = (sd(x)/sqrt(length(x))))
    })))
}

release_yr_coef = round(fixef(RES_full)[2] * scl, 1)
yrs_coef = round(fixef(RES_full)[3] * scl, 1)
only_release_coef = round(fixef(RES_rel_fit)[2]*scl, 1)

RES_CI <- round(posterior_interval(RES_full, pars = "yrs_in_trial") * scl, 1)
rel_CI = round(posterior_interval(RES_full, pars = "release_yr_c") * scl, 1)
only_release_CI = round(posterior_interval(RES_rel_fit, pars = "release_yr_c") * scl, 1)

m202_v_206 = round(yrs_coef * -19, -1)

id_est = ranef(RES_full)$id
id_est = cbind(id_est, posterior_interval(RES_full, regex_pars = "id:M"))
idx = order(id_est[,1], decreasing = FALSE)
id_est = id_est[idx,]

@

\section{Introduction}\label{intro}

\setlength{\parindent}{1cm}

Constraints on arable land are increasing simultaneous with the need
to increase total food production to meet a growing demand
\citep{Foley2011a, Godfray2010, Mueller2012}, which has necessitated
harvesting more grain per unit land area \citep{Lobell2009,
Tittonell2014}. Historically, agricultural research has been
successful in staving off the ``Malthusian catastrophe'' of demand
surpassing supply via continued yield improvement of staple
crops. However production systems are experiencing plateaus in grain
yield \citep{Grassini2013, Peng2010}. If this trend continues,
improvements per unit land area are no longer possible, and an
increase in the area under cultivation will be needed to meet food
demand, which carries with it undesirable ecological
implications. Therefore, it is of critical importance to better
understand why the rate of increase in grain yields has declined or
leveled off in intensified production systems.

California (CA) rice production represents one such production
system. Rice yields in CA during the period from 1920 -- 1990
increased rapidly (Fig. \ref{fig:state_yield}), in part due to
improvements in rice genetics (i.e., the ``Green Revolution'') through
the efforts of the California Cooperative Rice Research Foundation
(CCRRF; a collaboration between the University of California, the
USDA-Agricultural Research Service, and farmer-funded
research). However, since 1990 the rate of increase has slowed
(Fig. \ref{fig:state_yield}) despite continuous crop improvement
\citep{ccrrf}. For the past 30 years, each newly developed cultivar
has demonstrated a yield advantage over the previous cultivars in the
Statewide Variety Trials (an annual multi-site variety trial program
that compares experimental entries side-by-side with ``check''
cultivars). However, in a seeming contradiction these same reports
show stagnant yields across release announcements (i.e., reports
accompanying the official public release) from the previous years. For
example, when cultivar M-202 was released in 1985, it was reported to
yield an average of 11 Mg ha\tsups{-1} \citep[14\% moisture content
(MC);][]{m202}. In 2015, the reported average yield of the newest
released cultivar M-209 was 10.8 Mg ha\tsups{-1} \citep]14\%
MC;][]{m209}. In other rice systems, a decrease in a cultivar's yield
performance over time has been observed \citep{Peng1998, Peng2010,
Peng2000, Datta1995}, often termed yield erosion \citep{Peng1998}. It
has been speculated that this yield erosion is due to the inability of
a cultivar to adapt to changing biotic and abiotic conditions. To
better understand how plant improvement and yield erosion are
contributing to the improvement of rice yield in farmers' fields, we
teseted the hypotheses: either (1) the genetic yield advantage
increased over the last 30+ years, or (2) the genetic yield advantage
has not increased over the last 30+ because the erosion of yield over
time creates an illusion of improved yields.

\begin{figure}[H]
<<fig=TRUE, echo=FALSE>>=
CA_data = read.csv("../data/CA_rice_yield.csv", stringsAsFactors = FALSE)
CA_data$Value = as.numeric(gsub(",", "", CA_data$Value))/1000
CA_data = CA_data[order(CA_data$Year),]
plot(Value * 1.12 ~ Year, CA_data,
     # type = "l",
     pch = 16,
     xlab = "Year",
     ylab = expression("State avg. grain yield ( Mg ha"^-1~")"))
# mv_avg = filter(c(rep(mean(CA_data$Value[1:10] * 1.12), 5),
#                   CA_data$Value * 1.12,
#                   rep(mean(CA_data$Value[(nrow(CA_data) - 10):nrow(CA_data)] * 1.12), 5)),
#                 rep(1/11,11))
# lines(x = c(1907:1911, CA_data$Year, 2017:2021), y = mv_avg)

@
\caption{Average rice yields over the period \Sexpr{min(CA_data$Year)}
to \Sexpr{max(CA_data$Year)} for California, USA (US Dept. of
Agriculture National Agricultural Statistics Service, 2017). Rapid
gains in grain yield were experienced from approximately 1950 -- 1990,
but annual gains since 1990 have been slower.}
\label{fig:state_yield}
\end{figure}

\section{Methods}\label{methods}

\subsection{Site and data description}

The CCRRF Rice Experiment Station (RES) is located north-west of
Biggs, CA (39.4648, -121.7342), and has been the central location for
CCRRF's efforts to develop improved rice varieties adapted to CA since
1969. The climate at the RES is Mediterranean, characterized by mild
winters during which most of the annual precipitation occurs
(\Sexpr{18*24.5} to \Sexpr{25*24.5} mm annually) and warm summers
largely free of precipitation events. Soils at the RES are classified
as Esquon-Neerdobe clays with roughly 2\% soil organic matter
\citep{ssurgo} in the top 15cm. As the central location for the CCRRF
breeding program, the RES is researcher-managed (as opposed to
farmer-managed trials) and plants the majority of experimental plots for the
Statewide Variety Trials (approximately 3 to 4 times the number of
experimental plots compared to any other location). To avoid potential
complications due to site, climate, and management differences at the
other 6 to 8 farmer-managed trials per year we focused our
investigation on the RES.

As part of the annual Statewide Variety Trials, newly developed
entries are tested against officially released cultivars (checks) at
the RES in several trials spanning multiple planting dates per
year. For the purposes of this study, we concentrated on released
medium grain cultivars, as these cultivars constitute approximately
90\% of the planted rice area in CA. These cultivars are typcically
evaluated in the Statewide Variety Trials for 3 to 5 years prior to
their official release. For these cultivars planted in breeding trials
at the RES from \Sexpr{min(mod_data[,'year'])} --
\Sexpr{max(mod_data[,'year'])} plot level observations (replicated
3--4 times in a CRB design) were collected leading to a final data set
of \Sexpr{format(nrow(mod_data), big.mark = ',')} observations
representing \Sexpr{length(unique(mod_data[,'id']))} cultivars,
including every publicly released medium grain variety developed by
the CCRRF from \Sexpr{min(mod_data[,'release_yr'])} --
\Sexpr{max(mod_data[,'release_yr'])}. For each cultivar, the official
year of public release and the time elapsed (in years) since release was
calculated.

Experimental plots at the RES are managed similar to University of California
Cooperative Extension following prescribed best management
practice, including land preparation, fertility management, and pest
and disease control \citep{ucce}. For planting, rice seed is pre-germinated prior
to direct-seeding into pre-flooded fields. Plots were either 14.0 or
18.5 m\tsups{2}, and were harvested after physiological
maturity using a small-plot combine. All yields were converted to 14\%
moisture content prior to reporting.

\subsection{Statistical analysis}

To test each hypothesis, the influence of breeding and yield erosion
were quantified on yield over time. Yield was modeled as a response to
the number of years since the variety was officially released and the
year of release using a mixed-effects linear regression. To account
for similarities between years and between cultivars, year and
cultivar were included in the model as random effects.

\begin{equation}
\begin{split}
 yield_i = Intercept + (Rate_{yield\ erosion\ i}* {years\ since\
   release\ i}) \\+ (Rate_{yield\ improvement\ i}* {release\ year\ i}) 
 + \beta_{year\ i} + \beta_{cultivar\ i} \\
\end{split}
\end{equation}

Where $yield_i$ is the grain yield for plot $i$, the Intercept is the
average yield for the RES across varieties and years, $Rate_{yield\
erosion\ i}$ is rate of yield erosion per ${years\ since\ release\ i}$
(negative for years prior to release, zero at the year of release),
$Rate_{yield\ improvement\ i}$ is the rate of yield improvement per
year (${release\ year\ i}$), $\beta_{year}$ is the random effect of
year, and $\beta_{cultivar}$ is the random effect of cultivar. For
comparison to estimates where yield erosion is not taken into account,
the model was used as described above except omitting the $Rate_{yield\ erosion\ i}$ term.

% Add more detail to the methods section - tell how you get these
% numbers. "Evidence for erosion" - what does this mean for yield
% increases? Erosion 1st, 2nd implications of erosion. 
% Drop off -5 and -6

Data were processed and models were fit in R, an environment for
statistical computing \citep{R}. Models were fit using the `rstanarm'
package \citep{rstanarm}, an interface to Stan, a language for
probabilistic programming \citep{stan}. All model diagnostics,
including $\hat{R}$, effective size, and posterior predictive checks
were examined before reporting results. To test goodness of fit, the
full model was compared to models containing only the years since
release or the year of release using the `loo' package \citep{loo}, an
efficient means of conducting leave-one-out cross validation
\citep{loo_paper}. The 90\% credible interval (90\% CI), defined as
the interval containing 90\% of the distribution of estimated credible
parameter values given the data, was calculated as a measure of
uncertainty in the parameter estimates using the quantile method. The
90\% CI is more stable to perpetuation's and hence is preferable to the 95\%
interval \citep{rstanarm}. Complete data and code used for this
analysis is available at \url{github.com/mespe/rice_advances.git}.

\section{Results}\label{results}
% Invert this section, then talk about yield improvement accounting
% for yield erosion.

\begin{figure}[H]
<<fig=TRUE, width = 6, height = 5, out.width = '85%', echo=FALSE>>=

idx = mod_data[,"yrs_in_trial"] %in% -2:0

plot_data = prep_plot_data(mod_data[idx,"yield_kg"] / 1000 ,
                           mod_data[idx, "id"])

plot_data[,"yr"] = tapply(mod_data[idx, "release_yr"],
                          mod_data[idx,"id"], unique)

plot(mean ~ yr, data = plot_data,
     pch = 16,
     xlab = "Year of release",
     ylab = expression("Yield ( Mg ha"^-1~")"),
     ylim = c(9,13),
     xlim = c(1984, 2018.5))

segments(x0 = plot_data[,"yr"],
         y0 = plot_data[,"mean"] + plot_data[,"se"],
         y1 = plot_data[,"mean"] - plot_data[,"se"])

#Used locator() it interactively find label points 
aa = readRDS("f2_text_pos.rds")
text(x = aa$x, y = aa$y,
     labels = gsub("M", "M-", row.names(plot_data)),
     pos = 4, cex = 0.75)
#### Polygons
yrs <- 1975:2025

pp <- posterior_linpred(RES_full, re.form = NA,
                        newdata = data.frame(release_yr_c = yrs - 2000,
                                             yrs_in_trial = 0),
                        prob = 0.5)
pp <-  (pp * scl + center) / 1000
ints2 = apply(pp, 2, quantile, p = c(0.05, 0.95))

abline(a = ((((fixef(RES_full)[1] * scl) + center) ) -
           (2000 * release_yr_coef)) / 1000,
       b = release_yr_coef/1000,
       lty = 2)
polygon(x = c(yrs, rev(yrs)),
        y = c(ints2[1,],rev(ints2[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

# text(x = plot_data$yr, y = plot_data$mean,
#      labels = 1:nrow(plot_data), #gsub("M", "M-", row.names(plot_data)),
#      pos = 4, cex = 0.75)
# mm = lm(yield_kg/1000 ~ release_yr, data = mod_data[idx,])
# abline(mm, lty = 2)
# yrs = 1980:2020
# int = predict(mm, newdata = data.frame(release_yr = yrs), se.fit = TRUE)

# polygon(x = c(yrs, rev(yrs)),
#         y = c(int$fit + int$se.fit,rev(int$fit - int$se.fit)),
#         border = NA,
#         col = rgb(0,0,0,0.1))

@

\caption{The experimental yield of medium grain rice varieties for the
  two years prior to each cultivar's public release. Points
($\bullet$) and bars (|) represent the mean and standard error of the
raw data, respectively. The dashed line and shaded area are the
genetic yield improvement and credible interval as estimated by the model.}
\label{fig:first_yrs}
\end{figure}

\subsection{Genetic yield improvements}

Comparing the cultivars' yield during each's initial time (average 3 years) in the
experimental trials confirmed that genetic yield improvements have
been consistant over the 30 year study period
(Fig. \ref{fig:first_yrs}). As estimated by the model including both
yield erosion and yield improvement over time, there is
no conclusive evidence that the yields of medium grain rice cultivars
selected and released by CCRRF's breeding program have, on average,
increased over the last 30 years (Table \ref{tbl:coef}). The full
model which corrected for both the year of release and yield erosion
was estimated to have better out-of-sample predictive accuracy
compared to reduced models containing only one predictor (See
supplemental material for further details). The
estimated change in grain yield per year in the CCRRF's breeding
trials at the RES was estimated to be a negligible
\Sexpr{release_yr_coef} kg ha\tsups{-1} year\tsups{-1}, with both
small positive and negative trends within the credible interval (90\%
CI: \Sexpr{rel_CI[1]} to \Sexpr{rel_CI[2]})
(Fig. \ref{fig:first_yrs}).

When yield erosion was not corrected for in the analysis (i.e.,
omitted from the model), there is an apparent small increase per year
(approximately \Sexpr{only_release_coef} kg ha\tsups{-1}
year\tsups{-1}; 90\% CI \Sexpr{only_release_CI[1]} to
\Sexpr{only_release_CI[2]}) (Fig. \ref{fig:wo_erosion}). Although this
estimate is flawed (i.e., it ignores the confounding effect of yield
erosion over time and performed worse in out-of-sample predictive
accuracy), it confirms previous analyses of side-by-side comparisons
which concluded that newer cultivars out-preform older ones.

After correcting for yield erosion, some cultivars were higher
yielding compared to the overall average (as estimated by random
intercepts). The marginal estimate for the performance of a cultivar
averaged over the uncertainty in the other parameters in the model
(i.e., the random effect) shows three cultivars with higher than
average yields:
\Sexpr{paste(gsub('M','M-',rev(row.names(id_est))[c(1,2,4)]),collapse=',')}
(Figs. \ref{fig:first_yrs} \& \ref{fig:ids_comp}). A fourth, M-209,
might also have higher than average yield, though the CI for M-209
overlaps with zero (Fig. \ref{fig:ids_comp}).

\begin{figure}[H]

<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=
yrs <- 1975:2025

pp <- posterior_linpred(RES_rel_fit, re.form = NA,
                        newdata = data.frame(release_yr_c = yrs - 2000,
                                             yrs_in_trial = 0),
                        prob = 0.5)
pp <-  (pp * scl + center) / 1000
ints1 = apply(pp, 2, quantile, p = c(0.05, 0.95))

res = mod_data$site == "RES"
comp_yrs = 2011:2016
idx = mod_data$year %in% comp_yrs

plot_data = prep_plot_data(mod_data$yield_kg[idx] / 1000 ,
                           mod_data$id[idx])

plot_data$tmp = tapply(mod_data$release_yr[idx],
                       mod_data$id[idx], unique)

colnames(plot_data) = c("yield_kg", "yield_se", "release_yr")

# get corrected yield
wo_erosion = expand.grid(id = row.names(plot_data),
            year = comp_yrs,
            release_yr_c = 0,
            yrs_in_trial = 0)
wo_erosion$corr_yield = colMeans(posterior_predict(RES_full,
                               newdata = wo_erosion)) + center/1000
plot_data2 = prep_plot_data(wo_erosion$corr_yield,
                            wo_erosion$id)
plot_data2$release_yr = plot_data$release_yr

# par(mfrow = c(1,2),
#     mar = c(4,0,2,0.75),
#     oma = c(4,4,0.5,0.5))

plot(yield_kg ~ release_yr, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Release year",
     ylab = "", 
     ylim = c(8, 12.5),
     xaxt = "n",
     xlim = c(1980,2020))
axis(1, at = seq(1980, 2020, by = 10))
mtext(text = expression("Grain yield ( Mg ha"^-1~")"), side = 2, line = 2.5)
abline(a = ((fixef(RES_rel_fit)[1] * scl) + center) / 1000 -
      (2000 * fixef(RES_rel_fit)[2]),
       b = fixef(RES_rel_fit)[2],
       lty = 2)
polygon(x = c(yrs, rev(yrs)),
        y = c(ints1[1,],rev(ints1[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

with(plot_data, segments(x0 = release_yr, y0=yield_kg + yield_se,
                         y1 = yield_kg - yield_se))
text(x = plot_data$release_yr-0.4, y = plot_data$yield_kg + 0.07,
     labels = gsub("M", "M-", row.names(plot_data)),
     pos = 4, cex = 0.75)
text(x = 2000, y = 8, bquote(.(round(fixef(RES_rel_fit)[2] * scl, 1))~kg~ha^-1~yr^-1))
mtext("Without adjusting for yield erosion", adj = 0)

@
\caption{The relationship between yield and the release year of medium
grain rice cultivars for the last 6 years
(\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated without
correcting for yield erosion over time.}
\label{fig:wo_erosion}
\end{figure}

\begin{comment}
\begin{figure}[H]
<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=

plot(mean ~ release_yr, data = plot_data2, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Release year",
     ylab = "", #expression("Grain yield -corrected( Mg ha"^-1~")"),
     yaxt = "n",
     ylim = c(8, 12.5),
     xaxt = "n",
     xlim = c(1980,2020))
axis(1, at = seq(1980, 2020, by = 10))
abline(a = ((((fixef(RES_full)[1] * scl) + center) ) -
           (2000 * release_yr_coef)) / 1000,
       b = release_yr_coef/1000,
       lty = 2)
with(plot_data2, segments(x0 = release_yr, y0 = mean + se,
                         y1 = mean - se))
text(x = plot_data2$release_yr-0.4, y = plot_data2$mean + 0.07,
     labels = gsub("M", "M-", row.names(plot_data2)),
     pos = 4, cex = 0.75)
text(x = 2000, y = 8, bquote(.(release_yr_coef)~kg~ha^-1~yr^-1))
mtext("Adjusting for yield erosion", adj = 0)
polygon(x = c(yrs, rev(yrs)),
        y = c(ints2[1,],rev(ints2[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

@
% Change from a./b. to accounting for erosion, More details about how
% yield was corrected - 
\caption{The relationship between yield and the release year of medium
grain rice cultivars for the last 6 years
(\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated by 
correcting for yield erosion over time. Yield erosion was corrected
for by adjusting each observed yield by estimated amount lost 
  (\Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} after release).}
\label{fig:w_erosion}
\end{figure}
\end{comment}

\subsection{Yield erosion}
\begin{figure}[H]

<<fig=TRUE, echo=FALSE>>=
yrs_in <- -10:40
pp <- posterior_linpred(RES_full, re.form = NA,
                        newdata = data.frame(yrs_in_trial = yrs_in,
                                             release_yr_c = 0))
pp <-  (pp * scl + center) / 1000 
ints = apply(pp, 2, quantile, p = c(0.05, 0.95))

plot_data = prep_plot_data(mod_data$yield_kg / 1000,
                           mod_data$yrs_in_trial)

plot_data$tmp = as.numeric(row.names(plot_data))
colnames(plot_data) = c("yield_kg", "yield_se", "yrs_in_trial")

plot(yield_kg ~ yrs_in_trial, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Year since release",
     ylab = expression("Grain yield ( kg ha"^-1~")"),
     ylim = c(8.0, 12.5))
with(plot_data, segments(x0 = yrs_in_trial, y0=yield_kg + yield_se,
                         y1 = yield_kg - yield_se))
abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000 ,
       b = yrs_coef/1000)
polygon(x = c(yrs_in, rev(yrs_in)),
        y = c(ints[1,],rev(ints[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))
text(5, 8.0, bquote(.(yrs_coef)~kg~ha^-1~year^-1))

@
\caption{The relationship between yield and the years since release
  for medium grain rice varieties. The estimated erosion of yield over time
  (averaged over differences in cultivar and year) is estimated to be
  \Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1}. Points ($\bullet$)
  and bars (|) represent the mean and standard error of the raw data,
  respectively. The shaded area is the 90\% credible interval.}
% For each year, point size by n (open circles)
\label{fig:RES_year}
\end{figure}

\begin{figure}[H]
<<fig=TRUE, width = 8, height = 5, out.width = '85%', echo=FALSE>>=
tars <- c("M202","M206")
par(mfrow = c(1, length(tars)),
    mar = c(4,0,2,0),
    oma = c(4,4,0.5,0.5))

plot_data = lapply(tars, function(x) {
    dd = mod_data[mod_data$id == x & mod_data$site == "RES",]
    tmp = prep_plot_data(dd$yield_kg/1000 , dd$year)
    tmp$year = as.numeric(row.names(tmp))
    tmp
    })

invisible(sapply(1:2, function(i){
    plot(mean ~ year,
         data = plot_data[[i]],
         xlim = c(1985, 2016), ylim = c(7.5,13.5),
         pch = 16, main = gsub("M","M-", tars[i]),
         ylab = "",
         yaxt = "n",
         xlab = "Year",
         cex.axis = 0.8)
    segments(x0 = plot_data[[i]]$year,
             y0 = plot_data[[i]]$mean + plot_data[[i]]$se,
             y1 = plot_data[[i]]$mean - plot_data[[i]]$se)
    if(i == 1){
        axis(2, cex.axis = 0.8,)
        mtext(expression("Yield ( Mg ha"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }
    # abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000 -
    #            (2000 * yrs_coef/1000),
    #        b = yrs_coef/1000,
    #        lty = 1)
    lines(as.numeric(rownames(plot_data[[i]])),
          11.2 + (1:nrow(plot_data[[i]])) * (yrs_coef/1000))
    # abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000,
           # b = yrs_coef/1000)

    # abline(h = mean(mod_data$yield_kg[(mod_data$id == tars[i] &
    #                                    mod_data$site == "RES")] ),
    #        lty = 1, col = "grey50")
    # mod = lm(mean ~ year, data = plot_data[[i]])
    # tmp = predict(mod)
    # if(i == 1){
    #     abline(mod)
    # }else{
    #     lines(as.numeric(names(tmp)), tmp)
    # }
    abline(h = 11.2, lty = 2)
    # text(2000, 7.5, bquote(.(round(coef(mod)[2] * 1000, 1))~kg~ha^-1~year^-1))
}))

@
\caption{The performance of two major cultivars over the period
\Sexpr{min(mod_data[,'year'])} -- \Sexpr{max(mod_data[,'year'])}. The
solid line is the estimated yield erosion (\Sexpr{yrs_coef} kg
ha\tsups{-1} year\tsups{-1}), while the dashed line is the published
3-year average yield of M-202 at its release (1982-1984). Both
cultivars have a similar yield at their respective year of release,
and similar decreases over time.}
\label{fig:compare_yrs}
\end{figure}

The average decrease in yield over time once a cultivar is selected
(i.e., yield erosion) was estimated as \Sexpr{yrs_coef} kg
ha\tsups{-1} year\tsups{-1} (90\% CI: \Sexpr{RES_CI[2]} to
\Sexpr{RES_CI[1]}). The oldest cultivar in the trials had lost an
average of \Sexpr{round(yrs_coef*-36/1000,1)} Mg ha\tsups{-1} since
its release. The two most dominate medium grain rice cultivars, M-202
(widely planted from the late 1980s until the early 2000s) and M-206
(currently the most widely planted), experienced
similar trends in yield decline (Fig. \ref{fig:compare_yrs}; Table
\ref{tbl:coef}). The currently estimated
\Sexpr{m202_v_206} kg ha\tsups{-1} yield advantage of M-206 over M-202
can be attributed solely to the erosion of M-202's yield over time
(Figs \ref{fig:first_yrs} \& \ref{fig:compare_yrs}). Further
supporting this estimate of yield erosion, both cultivars yielded similarly at their
respective year of release (Fig. \ref{fig:compare_yrs}), despite the
fact that M-202 was released 19 years prior to M-206
(Fig. \ref{fig:compare_yrs}, Table \ref{tbl:ids}).

\section{Discussion}

We do not find conclusive evidence of increased genetic yield
potential in medium grain rice cultivars released by CCRRF's breeding
improvement program over the period \Sexpr{min(mod_data[,'year'])} to
\Sexpr{max(mod_data[,'year'])}. Since the estimated yield improvement
spans from \Sexpr{rel_CI[1]} to \Sexpr{rel_CI[2]} kg ha\tsups{-1}
year\tsups{-1}, both increases and decreases in the genetic yield
potential over the study period are plausible. However, even the
estimated yield improvement that does not account for yield erosion
(\Sexpr{only_release_coef} kg ha\tsups{-1} year\tsups{-1};
Fig. \ref{fig:wo_erosion}) is much lower than previous estimated,
i.e., 37 to 60 kg ha\tsups{-1} year\tsups{-1} \citep{Omar2014}. Based on the
estimates from this study, there is less than a 1\% probability that
the true value is higher than 45 kg ha\tsups{-1}
year\tsups{-1}. Therefore, we conclude that the CCRRF breeding program
has likely maintained but not substantially improved the yield
potential in the face of changing biotic and abiotic pressures.

We estimate that yield erosion has decreased the yield of medium grain
rice cultivars released by the CCRRF over the last 30+ years by an
average of \Sexpr{yrs_coef} Mg ha\tsups{-1}, potentially offsetting
any gains made by the crop improvement program. Biotic and abiotic forces have
previously been speculated to be responsible for yield erosion, since
the cultivar's genetics remain stable from the time of selection but
the conditions the cultivar may experience over time are continuously
changing \citep{Peng1998, Peng2000, Peng2010}. The trend observed here
is consistent with previous reports of the erosion of genetic gains of
cultivar IR8 at IRRI \citep{Peng1998, Peng2000}, suggesting the
erosion of yield over time is a phenomenon that widely affects rice
breeding programs. This observation highlights both the need for maintenance
of plant breeding and the difficulty of increasing genetic yield potential,
especially in systems such as CA where yield is approaching the
physiological limit \citep{espe16a}. Without maintenance breeding
(i.e., continuously selecting cultivars well adapted to current
conditions), CA rice production systems are predicted to have lost an
estimated \Sexpr{round(yrs_coef*(-36/1000),1)} Mg ha\tsups{-1} grain
yield over the study period.

However, a goal of many breeding programs is to not only maintain the
benchmark yield of the previous ``standard'' but to surpass it,
especially if we are to meet the target of 50\% increased food
production by 2050 \citep{Foley2011a}. Given the estimated yield
erosion, to hit an ambitious target of 0.5 Mg ha\tsups{-1}
(approximately 5\%) increase in yield per release (with roughly 4 to 5
years between releases), a new cultivar needs to yield an additional
\Sexpr{round(yrs_coef*-4,-1)} to \Sexpr{round(yrs_coef*-5,-1)} kg
ha\tsups{-1} in side-by-side comparison in addition to the 0.5 Mg
ha\tsups{-1} target improvement. Often side-by-side yield improvements
show an improvement in grain yield of 2 to 3\% per release, in which
case no absolute improvements in yield have been made beyond
maintaining yield. The evidence here is consistent with the
observation that the genetic yield potential has not increased
substantially in the last 30+ years (Fig. \ref{fig:first_yrs}).

\begin{figure}[H]
<<fig = TRUE, echo = FALSE, width=5, height=5>>=

plot(1, type = "n",
     # xaxt = "n",
     xlab = "Years",
     ylab = expression("Grain yield ( Mg ha"^-1~")"),
     ylim = c(8.5,11.5),
     xlim = c(0,40),
     xaxs = "i")

# arrows(12, 8.2, 27, 8.2, xpd = TRUE, lwd = 2)
abline(a = 10, b = fixef(RES_full)[3], lty = 2, lwd = 2)
text(25, 10.85, "(3) target above maintenance", cex = 0.8,
     srt = 18)#((atan(fixef(RES_full)[3] + 0.06) * 180))/pi)
abline(h = 10, lty = 3, lwd = 2)
text(25, 10.1, "(1) current situation", cex = 0.8)
abline(a = 10, b = fixef(RES_full)[3] + 0.06, lty = 4, lwd = 2)
text(25, 9.35, "(2) no maintenance breeding", cex = 0.8,
     srt = -18)

@ 
\caption{A conceptual diagram highlighting three scenarios: (1)
maintenance breeding, where losses over time are offset by
improvements via plant improvement, (2) no maintenance breeding, where
grain yield slowly erodes over time (as estimated from this study),
and (3) absolute improvement in grain yield (based on ``typical''
yield goals of 3\% increase per release), where gains in grain yield
exceed the losses over time.}
\label{fig:theory}
\end{figure}

California state average yields have increased slowly
despite there being little to no gains in the breeding trials
(Fig. \ref{fig:state_yield}). The evidence here suggests mechanisms
other than increased genetic yield potential are responsible for these gains
(Fig. \ref{fig:compare_yrs}). Improvement in agronomic management,
including more efficient nutrient management \citep{Lundy2012,
Lundy2015, linquist2009}, weed management \citep{Saito2008,
Pittelkow2012, linquist2008}, and other factors may all be contributing to
increases in the state average yield. However, these gains are far less
than those experienced during the preceeding era, and by their
nature agronomic gains cannot surpass the genetic yield potential
\citep{Fischer2015}. Judging from yields achieved at the RES
(Figs. \ref{fig:RES_year} \& \ref{fig:compare_yrs}), a
researcher-managed production system implementing best management
practices in more controlled situations compared to farmers' fields, state
average yields (approximately 10 Mg ha\tsups{-1} at the time of
writing; Fig. \ref{fig:state_yield}) may be approaching the current
genetic limit. Assuming that the level of management possible in small
experimental plots cannot be attained in farmers' fields and hence
experimental yields represent the attainable yield ceiling
\citep{Fischer2015}, agronomic gains in farmer's fields are nearing
their potential, consistent with physiological model-based estimates
of the attainable yield gap in CA rice production systems
\citep{espe16a}.

There are several cultivars that were found to have better than
average yields, even after correcting for yield erosion. However,
these cultivars (M-201, M-204, M-205, and possibly M-209;
Fig. \ref{fig:ids_comp})) are not widely adapted. It is known that
cool temperatures during the booting stage can have a large impact on
grain yields \citep{espe2017}, and the RES
experiences warmer temperatures than many other rice growing areas in
CA \citep{espe2017}. These cultivars may perform well at the RES but
not widely in CA. Therefore, it should be noted that while our
investigation has been concerned about increasing yield potential, the
estimates here might not capture yield stability. Yield stability is
notoriously difficult to quantify \citep[e.g.,][]{lobell2011}, hence
further investigation is needed to explore how yield stability has
changed over the study period.

To be clear, our investigation here is narrowly focused on the
rough-rice/field grain yield. Other advancements due to plant
improvement, including increased quality \citep{mckenzie1993} and
milling yield \citep{mckenzie1994}, have been made over the study
period and are important. Also, crop duration has been reduced. For
example, during this study period, M-202 and M-206 have been the
predominate varieties grown in CA, and while their yield potentials
were similar at the time of release (Fig. 5), M-206 matures about 6
days earlier than M-202 (Table 2). Shorter duration is very important
in CA due to an already short growing season and water limitations
(shorter duration varieties would presumably require less irrigation
water).

Planning to meet future
demand necessitates increases in grain yield \citep{Godfray2010,
Foley2011a, Mueller2012}. Meeting the challenge to produce more grain
without expansion of land under cultivation will rely on new genetics
with increased yield potential. The evidence here suggests that to
accomplish true increases over time, the improvement target needs take
into account decreases in the current ``standard'' cultivars' yield
over time (Fig. \ref{fig:theory}). While side-by-side comparisons will
continue to be useful, comparisons across years to each cultivar's
benchmark yield at the time of release should provide information on
the absolute improvements being made in yield potential over time. New
technologies may need to be adapted to further increase yield
potential, including hybrid rice.

\section{Conclusions}

Improvement in grain yield is essential to meeting the challenges of
feeding a growing population in a changing world. To achieve these
improvements, we discuss here that (1) continuous maintenance via plant
improvement is required, without which yields would decrease over
time, (2) the challenge of improving genetic yield potential is
substantial, since absolute gains require improvement in yield in
excess of the maintenance threshold, and (3) increasing yield through
improved management (agronomy) cannot exceed the ceiling of genetic
yield potential. Taken together, these conclusions underscore both the
necessity and substantial challenges facing rice breeding. We suggest
here that historical benchmarks be used in to evaluate promising rice
cultivars in addition to side-by-side comparisons to maintain
perspective of the absolute yield gains being made.

\newpage

\section{Tables}

<<echo=FALSE, results=tex>>=

CIs = posterior_interval(RES_full)[1:3,] * scl
CIs[1,] = round((CIs[1,] + center)/1000, 1)
CIs = round(CIs, 1)
med_eff = fixef(RES_full) * scl
med_eff[1] = round((med_eff[1] + center)/1000, 1)
med_eff = round(med_eff, 1)

RES_coef = cbind(med_eff, paste(CIs[,1],CIs[,2], sep = " to "))
rownames(RES_coef) = c("Intercept (Mg ha\\tsups{-1})",
                       "Yield improvement (kg ha\\tsups{-1} year\\tsups{-1})",
                       "Yield erosion (kg ha\\tsups{-1} year\\tsups{-1})")
colnames(RES_coef) = c("Median", "90\\% Credible interval")

print(xtable(RES_coef,
             caption = "The estimated median fixed effects and 90\\% credible intervals for the improvement in yield of medium grain rice cultivars developed by the California Cooperative Rice Research Foundation, California, USA over the time period 1984 -- 2016.",
             label = "tbl:coef",
             align = "lcc"),
      caption.placement="top",
      booktabs=TRUE,
      sanitize.colnames.function = function(x){x},
      sanitize.rownames.function = function(x){x})
@
\newpage

<<echo=FALSE, results=tex>>=
ids = aggregate(release_yr ~ id, data = mod_data, unique)
ids$dth = as.integer(round(aggregate(dth ~ id, data = mod_data, mean, na.rm = TRUE)$dth, 0))
colnames(ids) = c("Cultivar", "Release Year", "Days to 50% Heading")
ids = ids[order(ids[,2]),]
ids[,2] = as.integer(ids[,2])
ids[,1] = gsub("M", "M-", ids[,1])

print(xtable(ids,
             caption = "All named medium grain rice cultivars and their associated release year planted at the Rice Experiment Station (RES), Biggs, California from 1984 -- 2015.",
             label = "tbl:ids"),
      include.rownames = FALSE,
      booktabs = TRUE,
      caption.placement = "top")

@

% Notes: Duration has gotten shorter. Etc. Rearrange - End
% introduction with evidence of erosion, quatify the effect on yield
% increases.

\newpage

\section{References}\label{references}

\setlength{\parindent}{-0.2in} \setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}

\bibliography{breeding}

\newpage

\section{Supplemental tables and
figures}\label{supplemental-tables-and-figures}

\beginsupplement

\begin{figure}[H]

<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=

plot_data = lapply(tars, function(x) {
    dd = mod_data[mod_data$id == x & mod_data$site == "RES",]
    tmp = prep_plot_data(dd$yield_kg / 1000, dd$yrs_in_trial)
    tmp$yrs_in_trial = as.numeric(row.names(tmp))
    tmp
    })

plot(mean ~ yrs_in_trial,
     data = plot_data[[1]],
     xlim = c(-6, 32), ylim = c(6.5,14.5),
     pch = 16, #main = gsub("M","M-", tars[i]),
     ylab = expression("Yield ( kg ha"^-1~")"),
     xlab = "Years after release")
segments(x0 = plot_data[[1]]$yrs_in_trial,
         y0 = plot_data[[1]]$mean + plot_data[[1]]$se,
         y1 = plot_data[[1]]$mean - plot_data[[1]]$se)
points(mean ~ yrs_in_trial,
       data = plot_data[[2]],
       pch = 17, col = "grey40")
segments(x0 = plot_data[[2]]$yrs_in_trial,
         y0 = plot_data[[2]]$mean + plot_data[[2]]$se,
         y1 = plot_data[[2]]$mean - plot_data[[2]]$se,
         col = "grey40")
abline(lm(yield_kg/1000 ~ yrs_in_trial,
          data = mod_data[(mod_data$id %in% tars & mod_data$site == "RES"),]),
       lty = 1, col = "grey50")
legend("topright", legend = gsub("M","M-", tars),
       pch = c(16, 17), col = c("black","grey40"))

@
\caption{The performance of two major cultivars at the CA Rice
  Experiment Station (RES) since the year each cultivar was
  released. Both M-202 ($\bullet$) and M-206 ($\blacktriangle$) have similar
  intercepts, suggesting similar yield at the time of release and an
  apparent yield advantage of the newer cultivar, M-206, might be
  explained by a yield loss of M-202 over time. Points represent the
  mean, and bars (|) the standard error.}
\label{fig:compare_since_release}
\end{figure}

\newpage

% \begin{figure}[H]
% <<fig=TRUE, echo=FALSE>>=
% tars <- c("M202","M206")
% comp_yrs = 2004:2005

% boxplot(yield_kg/1000 ~ id, data = mod_data[(mod_data$id %in% tars) &
%                                             (mod_data$year %in% comp_yrs),],
%         ylab = expression("Grain yield ( Mg ha"^-1~")"))

% @
% \caption{An example of a typical side-by-side comparison between a older (M-202)
%   and newer (M-206) cultivar. Shown is the performance of these cultivars at the CA Rice
%   Experiment Station (RES) over the period \Sexpr{min(comp_yrs)} --
%   \Sexpr{max(comp_yrs)}. The newer
%   cultivar, M-206, appears to have a yield advantage over the older
%   cultivar, M-202, though this comparison does not account for the 10
%   year difference in years since release between the cultivars.}
% \label{fig:compare_vars}
% \end{figure}
% \newpage

\begin{figure}[H]
<<echo = FALSE, fig = TRUE, width = 7, height = 6>>=
library(ggplot2)
library(ggthemes)
ggplot(mod_data, aes(x = year, y = yield_kg/1000)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_smooth(method = "lm", color = "black") +
    theme_few() +
    xlab("Year") +
    ylab(expression("Grain yield ( Mg ha"^-1~")")) + 
    facet_wrap(~id, ncol = 4)

@
\caption{The yield of all medium grain released cultivars at the
  California Cooperative Rice Research Foundation Rice Experiment
  Station over the period \Sexpr{min(mod_data[,'year'])} --
  \Sexpr{max(mod_data[,'year'])}.}
\label{fig:all_cult}
\end{figure}

\newpage
\begin{figure}[ht]
<<echo=FALSE, fig = TRUE>>=

dotchart(id_est[,1],
         labels = gsub("M", "M-", row.names(id_est)),
         pch = 16,
         xlim = c(min(id_est[,2]),max(id_est[,3])),
         xlab = expression("Estimated difference from average ( Mg ha"^-1~")"))
segments(x0 = id_est[,2], x1 = id_est[,3], y0 = 1:nrow(id_est))
abline(v = 0, lty = 2)
@
\caption{The marginal estimates for the difference between cultivars,
  averaged over the uncertainty in the fixed effect parameters (i.e.,
  yield erosion and annual yield gains).}
\label{fig:ids_comp}
\end{figure}

\newpage

\end{document}