### 2017-10-01 output/freq.Rmd
Adds secondary analysis

---
output: pdf_document
title: "Supplemental material"
date: "Oct. 1, 2017"
geometry: margin=1in
---

```{r setup, echo = FALSE}
load("../data/model_data.rda")
library(lme4)

```

This document replicates the analysis in "Rice yield improvements
through plant breeding are offset by inherent yield declines over
time" (Espe et al, 2017) using classical/frequentist statistics to
verify the results are consistent regardless of statistical method.

The complete data, code, and manuscript is available from \url{https://github.com/mespe/rice_advances}.

# Model excluding yield erosion

```{r}
no_erosion = lmer(yield_kg ~ release_yr +
                      (1|id) + (1|year), data = mod_data)

summary(no_erosion)
confint(no_erosion, parm = c("release_yr"),
        method = "boot", level = 0.9)
```

# Model including yield erosion

```{r}
erosion = lmer(yield_kg ~ release_yr + yrs_in_trial +
                   (1|id) + (1|year), data = mod_data)

summary(erosion)
confint(erosion, parm = c("release_yr","yrs_in_trial"),
        method = "boot", level = 0.9)
```