### 2017-10-29 output/breed_paper.Rnw
Updated with Bruce's comments - need to revise.

\documentclass[12pt,]{elsarticle}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{longtable,booktabs}
\usepackage{comment}
\usepackage{graphicx,grffile, textcomp}
\usepackage{setspace}
\usepackage{etoolbox}
\usepackage{float}
\doublespacing
\usepackage{lineno}
\usepackage{lscape} %% Harvard
\bibliographystyle{model2-names}\biboptions{authoryear}
\pagenumbering{gobble}
%% For figure numbers in supplementary info
%% from http://bytesizebio.net/2013/03/11/adding-supplementary-tables-and-figures-in-latex/
\newcommand{\beginsupplement}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
}

\newcommand{\tsups}{\textsuperscript}
\usepackage{Sweave}
\SweaveOpts{prefix.string=graphics/report, png=TRUE}

\begin{document}

\begin{frontmatter}

\title{Rice yield improvements through plant breeding are offset by
  inherent yield declines over time}
\author[ucd]{Matthew B. Espe\corref{corresauthor}}
\cortext[corresauthor]{Corresponding author} \ead{mespe@ucdavis.edu}
\author[ucd]{David Mackill}
\author[ucd]{Jim E. Hill}
\author[ucce2]{Michelle Leinfelder-Miles}
\author[ucce]{Luis A. Espino}
\author[ucce]{Randall Mutters}
\author[ucd]{Chris van Kessel}
\author[ucd]{Bruce A. Linquist}

\address[ucd]{Dept. of Plant Sciences, University of California --
Davis, Davis CA 95616, USA}

\address[ucce]{University of California Cooperative Extension,
Oroville, CA 95965, USA}

\address[ucce]{University of California Cooperative Extension,
Colusa, CA 95932, USA}

\address[ucce2]{University of California Cooperative Extension,
Stockton, CA 95206, USA}

\begin{abstract}

Meeting the challenge of feeding a growing population with
increasingly limited resources will require increasing the yield
potential of staple crops, such as rice. Yet many high-yielding,
intensive production systems have experienced slow rates of yield
improvement in recent years despite a demonstrated increase in the
yield potential of new crop cultivars. We analyzed experimental data
from one such cropping system, i.e., California (CA) rice, in order to
quantify improvements made in the genetic yield potential obtained
through plant breeding. California rice systems are among the highest
in the world and close to maximum yield potential. Specifically, the
hypothesis was tested that if rice cultivar yields decline over time
(termed yield erosion) then apparent yield increases in side-by-side
yield comparison tests will not reflect increases in yield
potential. This hypothesis was tested using experimental yield data
over 30+ years at the California Cooperative Rice Research Foundation
Rice Experiment Station. Results show an apparent increase in yield in
side-by-side comparisons between old and new rice varieties was
observed when the yield erosion was not considered. However, the
yields of older varieties eroded at an estimated rate of 29.3 kg
ha\tsups{-1} year\tsups{-1} (90\% credible interval -4.4 to -53.3)
after initial selection. Once this effect was considered, the yield
advantage of newer cultivars over old was uncertain (-3.3 kg
ha\tsups{-1} year\tsups{-1}, 90\% credible interval -36.1 to
31.5). These results highlight the importance of continuous crop
improvement and deployment of new cultivars as yield will decline over
time without it. To increase the genetic yield potential of rice in
the face of this phenomenon, higher yield targets are needed to
counter-act the decrease over time in the current cultivars.

\end{abstract}

\begin{keyword} rice, yield potential, breeding, yield erosion
\end{keyword}

\end{frontmatter}

\newpage

\textbf{Abbreviations}

CA: California

CI: Credible Interval

RES: Rice Experiment Station

CCRRF: California Cooperative Rice Research Foundation

\newpage

%\linenumbers

<<echo=FALSE>>=

library(rstan)
library(rstanarm)
library(lme4)
library(xtable)

load("../data/model_data.rda")
load("h1_fit.Rda")
load("bivar_fit.Rda")
load("../data/all_vt_weather.Rda")

prep_plot_data = function(x, idx)
{
    as.data.frame(do.call(rbind, tapply(x, idx, function(x){
        c(mean = mean(x), se = (sd(x)/sqrt(length(x))))
    })))
}

release_yr_coef = round(fixef(RES_full)[2] * scl, 1)
yrs_coef = round(fixef(RES_full)[3] * scl, 1)
only_release_coef = round(fixef(RES_rel_fit)[2]*scl, 1)

RES_CI <- round(posterior_interval(RES_full, pars = "yrs_in_trial") * scl, 1)
rel_CI = round(posterior_interval(RES_full, pars = "release_yr_c") * scl, 1)
only_release_CI = round(posterior_interval(RES_rel_fit, pars = "release_yr_c") * scl, 1)

m202_v_206 = round(yrs_coef * -19, -1)

id_est = ranef(RES_full)$id
id_est = cbind(id_est, posterior_interval(RES_full, regex_pars = "id:M"))
idx = order(id_est[,1], decreasing = FALSE)
id_est = id_est[idx,]

@

\section{Introduction}\label{intro}

\setlength{\parindent}{1cm}

Constraints on arable land are increasing simultaneous with the need
to increase total food production to meet a growing demand
\citep{Foley2011a, Godfray2010, Mueller2012}, which has necessitated
harvesting more grain per unit land area \citep{Lobell2009,
Tittonell2014}. Historically, agricultural research has been
successful in staving off the ``Malthusian catastrophe'' of demand
surpassing supply via continued yield improvement of staple
crops. However, many production systems are experiencing plateaus in
grain yield \citep{Grassini2013, Peng2010}. If this trend continues,
improvements per unit land area are no longer possible, and an
increase in the area under cultivation will be needed to meet food
demand, which carries with it undesirable ecological
implications~\citep{Foley2011a, tilman1999}. Therefore, it is of
critical importance to better understand why the rate of increase in
grain yields has declined or leveled off in intensified production
systems.

California (CA) rice represents one such production system. Rice is
grown primarily in the Sacramento Valley which is characterized by
having a Mediterranean climate, with long days, dry growing season,
and relatively free of pests and diseases, which lead to some of the
highest yields in the world~\citep{FAO2016}. Production in CA is
predominately focused on premium quality medium grain
\textit{japonica} cultivars (e.g., CalRose rice), and rice from CA is
recognized globally for its quality
(\url{http://agfax.com/2015/11/03/rice-calrose-wins-best-rice-world-competition/}). Most
cultivars in use in CA are developed by the California Cooperative
Rice Research Foundation (CCRRF; a collaboration between the
University of California, the USDA-Agricultural Research Service, and
farmer-funded research). In part due to improvements in rice genetics,
rice yields in CA increased rapidly during the period from 1920 --
1990; however, since the 1990s, the rate of yield increase has slowed
(Fig. \ref{fig:state_yield}) despite continuous crop
improvement. Based on a yield-gap analysis, on-farm yields in the
major rice growing region of CA is 73 to 76\% of maximum yield
potential~\citep{espe16a}. In highly intensive systems such as this,
\citet{grassini2011} have shown that farmers are capable of attaining
85\% of the maximum yield potential. Thus, given that farmers are near
the attainable yield potential, increases in yield are likely to be
slow and \citet{espe16a} reported that indeed this was the case
with yields on average increasing by about 50 to 62 kg ha\tsups{-1} between
1999 and 2014.

Certainly one question in looking at this situation is, ``do new
varieties have increased yield potential''?  Like many breeding
programs, one of the goals of the CCRRF is to increase yields through
breeding varieties with increased yield potential. However, when the
popular cultivar M-202 was released in 1985, it was reported to yield
an average of 11 Mg ha\tsups{-1} \citep[14\% moisture content
(MC);][]{m202}. In 2015, the reported average yield of the newest
released cultivar M-209 was 10.8 Mg ha\tsups{-1} \citep[14\%
MC;][]{m209}. However, in annual breeding reports and varietal release
announcements, based on yield performance in side-by-side variety
trails, data show that new varieties usually out yield the older
varieties. In other rice systems, a decrease in a cultivar's yield
performance over time has been observed \citep{Peng1998, Peng2010,
Peng2000, Datta1995}, often termed yield erosion \citep{Peng1998}. It
has been speculated that this yield erosion is due to the inability of
a cultivar to adapt to changing biotic and abiotic
conditions~\citep{Peng1998, Peng2010, Peng2000, Datta1995}. Yield
erosion may explain the apparent contradiction above, in which
side-by-side yield comparisons show yield improvements in varieties
while overall there may be little change in yield potential. To
address this issue we tested the hypothesis that if rice cultivar
yields erode over time then apparent yield increases in side-by-side
yield comparison tests will not reflect true yield increases.o

\begin{figure}[H]
<<fig=TRUE, echo=FALSE>>=
CA_data = read.csv("../data/CA_rice_yield.csv", stringsAsFactors = FALSE)
CA_data$Value = as.numeric(gsub(",", "", CA_data$Value))/1000
CA_data = CA_data[order(CA_data$Year),]
plot(Value * 1.12 ~ Year, CA_data,
     # type = "l",
     pch = 16,
     xlab = "Year",
     ylab = expression("State avg. grain yield ( Mg ha"^-1~")"))
# mv_avg = filter(c(rep(mean(CA_data$Value[1:10] * 1.12), 5),
#                   CA_data$Value * 1.12,
#                   rep(mean(CA_data$Value[(nrow(CA_data) - 10):nrow(CA_data)] * 1.12), 5)),
#                 rep(1/11,11))
# lines(x = c(1907:1911, CA_data$Year, 2017:2021), y = mv_avg)

@
\caption{Average rice yields over the period \Sexpr{min(CA_data$Year)}
to \Sexpr{max(CA_data$Year)} for California, USA (US Dept. of
Agriculture National Agricultural Statistics Service, 2017). Rapid
gains in grain yield were experienced from approximately 1950 -- 1990,
but annual gains since 1990 have been slower.}
\label{fig:state_yield}
\end{figure}

\section{Methods}\label{methods}

\subsection{Site and data description}

The CCRRF Rice Experiment Station (RES) is located near Biggs, CA
(39.4648, -121.7342), and has been the central location for CCRRF's
efforts to develop improved rice cultivars adapted to CA since
1969. The California Statewide Variety Trails evaluates current and
promising varieties at the RES and six to eight on-farm trials around
the state each year. To avoid potential complications due to site,
climate, and management differences at the farmer-managed trials per
year, we focused our investigation on the RES. The RES is
researcher-managed and plants the majority of the experimental plots
for the statewide program (approximately 3 to 4 times the number of
experimental plots compared to other on-farm locations). The climate
at the RES is Mediterranean, characterized by mild winters during
which most of the annual precipitation occurs (\Sexpr{18*24.5} to
\Sexpr{25*24.5} mm annually) and warm summers largely free of
precipitation events. Soils at the RES are classified as
Esquon-Neerdobe clays with roughly 2\% soil organic matter
\citep{ssurgo} in the top 15cm.

As part of the annual Statewide Variety Trials, newly developed
entries are tested against officially released cultivars (checks) at
the RES in several trials spanning multiple planting dates per
year. For the purposes of this study, we concentrated on released
medium grain cultivars, as these cultivars constitute approximately
90\% of the planted rice area in CA. Newly developed entries are typically
evaluated in these trials for 3 to 5 years prior to
their official release. For these cultivars planted in breeding trials
at the RES from \Sexpr{min(mod_data[,'year'])} --
\Sexpr{max(mod_data[,'year'])} plot level observations (replicated
3--4 times in a CRB design) were collected leading to a final data set
of \Sexpr{format(nrow(mod_data), big.mark = ',')} observations
representing \Sexpr{length(unique(mod_data[,'id']))} cultivars and
included every publicly released medium grain cultivar developed by
the CCRRF from \Sexpr{min(mod_data[,'release_yr'])} --
\Sexpr{max(mod_data[,'release_yr'])}. For each cultivar, the official
year of public release and the time elapsed (in years) since release was
determined.

Experimental plots at the RES are managed similar to University of California
Cooperative Extension following prescribed best management
practice for land preparation, fertility rates, and pest
and disease control \citep{ucce}. As is typical for water-seeded
systems, for planting, rice seed is pre-germinated prior
to direct-seeding into pre-flooded fields. Plots were either 14.0 or
18.5 m\tsups{2}, and were harvested after physiological
maturity using a small-plot combine. All yields were converted to 14\%
moisture content prior to reporting.

\subsection{Statistical analysis}

To test the hypothesis, the influence of breeding and yield erosion
were quantified on yield over time using a Bayesian hierarchical
model. Yield was modeled as a response to the number of years since
the cultivar was officially released and the year of release using a
mixed-effects linear regression. To account for similarities between
years and between cultivars, year and cultivar were included in the
model as random effects.

\begin{equation}
\begin{split}
 yield_i = Intercept + (Rate_{yield\ erosion\ i}* {years\ since\
   release\ i}) \\+ (Rate_{yield\ improvement\ i}* {release\ year\ i}) 
 + \beta_{year\ i} + \beta_{cultivar\ i} \\
\end{split}
\end{equation}

Where $yield_i$ is the grain yield for plot $i$, the Intercept is the
average yield for the RES across cultivars and years, $Rate_{yield\
erosion\ i}$ is rate of yield erosion per ${years\ since\ release\ i}$
(negative for years prior to release, zero at the year of release),
$Rate_{yield\ improvement\ i}$ is the rate of yield improvement per
year (${release\ year\ i}$), $\beta_{year}$ is the random effect of
year, and $\beta_{cultivar}$ is the random effect of cultivar. For
comparison to estimates where yield erosion is not taken into account,
the model was used as described above except omitting the $Rate_{yield\ erosion\ i}$ term.

% Add more detail to the methods section - tell how you get these
% numbers. "Evidence for erosion" - what does this mean for yield
% increases? Erosion 1st, 2nd implications of erosion. 
% Drop off -5 and -6

Data were processed and models were fit in R, an environment for
statistical computing \citep{R}. Models were fit using the `rstanarm'
package \citep{rstanarm}, an interface to Stan, a language for
probabilistic programming \citep{stan}. All model diagnostics,
including $\hat{R}$, effective size, and posterior predictive checks
were examined before reporting results. To test goodness of fit, the
full model was compared to models containing only the years since
release or the year of release using the `loo' package \citep{loo}, an
efficient means of conducting leave-one-out cross validation
\citep{loo_paper}. The 90\% credible interval (90\% CI), defined as
the interval containing 90\% of the distribution of estimated credible
parameter values given the data, was calculated as a measure of
uncertainty in the parameter estimates using the quantile method. The
90\% CI is more stable to perpetuation's and hence is preferable to the 95\%
interval \citep{rstanarm}. To verify the estimates, the analysis was
also conducted using classical methods (frequentist mixed-effects
model fit by maximum likelihood; See supplemental material for
details). Complete data and code used for this
analysis is available at \url{github.com/mespe/rice_advances.git}.

\section{Results}\label{results}
% Invert this section, then talk about yield improvement accounting
% for yield erosion.

\begin{figure}[H]
<<fig=TRUE, width = 6, height = 5, out.width = '85%', echo=FALSE>>=

idx = mod_data[,"yrs_in_trial"] %in% -2:0

plot_data = prep_plot_data(mod_data[idx,"yield_kg"] / 1000 ,
                           mod_data[idx, "id"])

plot_data[,"yr"] = tapply(mod_data[idx, "release_yr"],
                          mod_data[idx,"id"], unique)

plot(mean ~ yr, data = plot_data,
     pch = 16,
     xlab = "Year of release",
     ylab = expression("Yield ( Mg ha"^-1~")"),
     ylim = c(9,13),
     xlim = c(1984, 2018.5))

segments(x0 = plot_data[,"yr"],
         y0 = plot_data[,"mean"] + plot_data[,"se"],
         y1 = plot_data[,"mean"] - plot_data[,"se"])

#Used locator() it interactively find label points 
aa = readRDS("f2_text_pos.rds")
text(x = aa$x, y = aa$y,
     labels = gsub("M", "M-", row.names(plot_data)),
     pos = 4, cex = 0.75)
#### Polygons
yrs <- 1975:2025

pp <- posterior_linpred(RES_full, re.form = NA,
                        newdata = data.frame(release_yr_c = yrs - 2000,
                                             yrs_in_trial = 0),
                        prob = 0.5)
pp <-  (pp * scl + center) / 1000
ints2 = apply(pp, 2, quantile, p = c(0.05, 0.95))

abline(a = ((((fixef(RES_full)[1] * scl) + center) ) -
           (2000 * release_yr_coef)) / 1000,
       b = release_yr_coef/1000,
       lty = 2)
polygon(x = c(yrs, rev(yrs)),
        y = c(ints2[1,],rev(ints2[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

# text(x = plot_data$yr, y = plot_data$mean,
#      labels = 1:nrow(plot_data), #gsub("M", "M-", row.names(plot_data)),
#      pos = 4, cex = 0.75)
# mm = lm(yield_kg/1000 ~ release_yr, data = mod_data[idx,])
# abline(mm, lty = 2)
# yrs = 1980:2020
# int = predict(mm, newdata = data.frame(release_yr = yrs), se.fit = TRUE)

# polygon(x = c(yrs, rev(yrs)),
#         y = c(int$fit + int$se.fit,rev(int$fit - int$se.fit)),
#         border = NA,
#         col = rgb(0,0,0,0.1))

@

\caption{The experimental yield of medium grain rice cultivars for the
  two years prior to each cultivar's public release. Points
($\bullet$) and bars (|) represent the mean and standard error of the
raw data, respectively. The dashed line and shaded area are the
genetic yield improvement and credible interval as estimated by the model.}
\label{fig:first_yrs}
\end{figure}

\subsection{Cultivar yields at time of release}

There was agreement between the estimates from the Bayesian and
classical approaches.  The Bayesian estimates are presented here,
though the classical estimates can be found in the supplemental
material.  Comparing the cultivars' yield during each's initial time
(average 3 years) in the experimental trials just before being
released showed that across cultivars, yields ranged from
\Sexpr{round(plot_data["M207", "mean"], 2)} (M-207) to
\Sexpr{round(plot_data["M205", "mean"], 2)} Mg ha\tsups{-1}
(M-205). Evaluating these initial yields over time indicates there was
no evidence of genetic yield improvements during the 30-year study
period (Fig. 2). Individual exceptions to this were two varieties
(M-204 and M-205) that had higher yields than the others.

\subsection{Yield erosion}
\begin{figure}[H]

<<fig=TRUE, echo=FALSE>>=
yrs_in <- -10:40
pp <- posterior_linpred(RES_full, re.form = NA,
                        newdata = data.frame(yrs_in_trial = yrs_in,
                                             release_yr_c = 0))
pp <-  (pp * scl + center) / 1000 
ints = apply(pp, 2, quantile, p = c(0.05, 0.95))

plot_data = prep_plot_data(mod_data$yield_kg / 1000,
                           mod_data$yrs_in_trial)

plot_data$tmp = as.numeric(row.names(plot_data))
colnames(plot_data) = c("yield_kg", "yield_se", "yrs_in_trial")

plot(yield_kg ~ yrs_in_trial, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Year since release",
     ylab = expression("Grain yield ( kg ha"^-1~")"),
     ylim = c(8.0, 12.5))
with(plot_data, segments(x0 = yrs_in_trial, y0=yield_kg + yield_se,
                         y1 = yield_kg - yield_se))
abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000 ,
       b = yrs_coef/1000)
polygon(x = c(yrs_in, rev(yrs_in)),
        y = c(ints[1,],rev(ints[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))
text(5, 8.0, bquote(.(yrs_coef)~kg~ha^-1~year^-1))
mtext("Adjusting for yield erosion", adj = 0)

@
\caption{The relationship between yield and the years since release
  for medium grain rice cultivars. Negative values on the X axis
  indicate experimental yields prior to the official release. The
  estimated erosion of yield over time (averaged over differences in
  cultivar and year) is estimated to be \Sexpr{yrs_coef} kg ha\tsups{-1}
  year\tsups{-1}. Points ($\bullet$) and bars (|) represent the mean and
  standard error of the raw data, respectively. The shaded area is the
  90\% credible interval.}
\label{fig:RES_year}
\end{figure}

The yields of each cultivar were evaluated over time relative to the
time when the cultivar was released using the full model which
accounts for both yield erosion and yield improvements over time. The
average decrease in yield over time (i.e., yield erosion) once a
cultivar was selected was estimated \Sexpr{yrs_coef} kg ha\tsups{-1}
year\tsups{-1} (90\% CI: \Sexpr{RES_CI[2]} to \Sexpr{RES_CI[1]}) (Fig
\ref{fig:RES_year}; Table \ref{tbl:coef}). The oldest cultivar in the
trials (\Sexpr{unique(mod_data[which.min(mod_data[,"year"]),"id"])}
had lost an average of \Sexpr{round(yrs_coef*-36/1000,1)} Mg
ha\tsups{-1} since its release. The most dominate medium grain rice
cultivar, M-202 (widely planted from the late 1980s until the early
2000s) experienced similar trends in yield decline
(Fig. \ref{fig:RES_year}; Table \ref{tbl:coef}). M-206 replaced M-202
as the most widely grown cultivar in CA and the current estimated 560
kg ha-1 yield advantage of M-206 over M-202 can be attributed solely
to the erosion of M-202's yield over time (Figs \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}). Further supporting this estimate of yield
erosion, both cultivars yielded similarly at their respective year of
release (Fig. \ref{fig:compare_yrs}), despite the fact that M-202 was
released 19 years prior to M-206 (Fig. \ref{fig:compare_yrs}, Table
\ref{tbl:ids}).

When yield erosion was not corrected for in the analysis (i.e.,
omitted from the model), there is an apparent small increase per year
(approximately \Sexpr{only_release_coef} kg ha\tsups{-1}
year\tsups{-1}; 90\% CI \Sexpr{only_release_CI[1]} to
\Sexpr{only_release_CI[2]}) (Fig. \ref{fig:wo_erosion}).  However,
using the full model which corrected for both the year of release and
yield erosion, indicates that the estimated change in grain yield per
year in the CCRRF's breeding trials at the RES was
\Sexpr{release_yr_coef} kg ha\tsups{-1} year\tsups{-1}, with both
small positive and negative trends within the credible interval (90\%
CI: \Sexpr{rel_CI[1]} to \Sexpr{rel_CI[2]})
(Fig. \ref{fig:first_yrs}).

After correcting for yield erosion, some cultivars were higher
yielding compared to the overall average (as estimated by random
intercepts). The marginal estimate for the performance of a cultivar
averaged over the uncertainty in the other parameters in the model
(i.e., the random effect) shows three cultivars with higher than
average yields:
\Sexpr{paste(gsub('M','M-',rev(row.names(id_est))[c(1,2,4)]),collapse=',')}
(Figs. \ref{fig:first_yrs} \& \ref{fig:ids_comp}).

\begin{figure}[H]

<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=
yrs <- 1975:2025

pp <- posterior_linpred(RES_rel_fit, re.form = NA,
                        newdata = data.frame(release_yr_c = yrs - 2000,
                                             yrs_in_trial = 0),
                        prob = 0.5)
pp <-  (pp * scl + center) / 1000
ints1 = apply(pp, 2, quantile, p = c(0.05, 0.95))

res = mod_data$site == "RES"
comp_yrs = 2011:2016
idx = mod_data$year %in% comp_yrs

plot_data = prep_plot_data(mod_data$yield_kg[idx] / 1000 ,
                           mod_data$id[idx])

plot_data$tmp = tapply(mod_data$release_yr[idx],
                       mod_data$id[idx], unique)

colnames(plot_data) = c("yield_kg", "yield_se", "release_yr")

# get corrected yield
wo_erosion = expand.grid(id = row.names(plot_data),
            year = comp_yrs,
            release_yr_c = 0,
            yrs_in_trial = 0)
wo_erosion$corr_yield = colMeans(posterior_predict(RES_full,
                               newdata = wo_erosion)) + center/1000
plot_data2 = prep_plot_data(wo_erosion$corr_yield,
                            wo_erosion$id)
plot_data2$release_yr = plot_data$release_yr

# par(mfrow = c(1,2),
#     mar = c(4,0,2,0.75),
#     oma = c(4,4,0.5,0.5))

plot(yield_kg ~ release_yr, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Release year",
     ylab = "", 
     ylim = c(8, 12.5),
     xaxt = "n",
     xlim = c(1980,2020))
axis(1, at = seq(1980, 2020, by = 10))
mtext(text = expression("Grain yield ( Mg ha"^-1~")"), side = 2, line = 2.5)
abline(a = ((fixef(RES_rel_fit)[1] * scl) + center) / 1000 -
      (2000 * fixef(RES_rel_fit)[2]),
       b = fixef(RES_rel_fit)[2],
       lty = 2)
polygon(x = c(yrs, rev(yrs)),
        y = c(ints1[1,],rev(ints1[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

with(plot_data, segments(x0 = release_yr, y0=yield_kg + yield_se,
                         y1 = yield_kg - yield_se))
text(x = plot_data$release_yr-0.4, y = plot_data$yield_kg + 0.07,
     labels = gsub("M", "M-", row.names(plot_data)),
     pos = 4, cex = 0.75)
text(x = 2000, y = 8, bquote(.(round(fixef(RES_rel_fit)[2] * scl, 1))~kg~ha^-1~yr^-1))
mtext("Without adjusting for yield erosion", adj = 0)

@
\caption{The relationship between yield and the release year of medium
grain rice cultivars for the last 6 years
(\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated without
correcting for yield erosion over time.}
\label{fig:wo_erosion}
\end{figure}

\begin{comment}
\begin{figure}[H]
<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=

plot(mean ~ release_yr, data = plot_data2, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Release year",
     ylab = "", #expression("Grain yield -corrected( Mg ha"^-1~")"),
     yaxt = "n",
     ylim = c(8, 12.5),
     xaxt = "n",
     xlim = c(1980,2020))
axis(1, at = seq(1980, 2020, by = 10))
abline(a = ((((fixef(RES_full)[1] * scl) + center) ) -
           (2000 * release_yr_coef)) / 1000,
       b = release_yr_coef/1000,
       lty = 2)
with(plot_data2, segments(x0 = release_yr, y0 = mean + se,
                         y1 = mean - se))
text(x = plot_data2$release_yr-0.4, y = plot_data2$mean + 0.07,
     labels = gsub("M", "M-", row.names(plot_data2)),
     pos = 4, cex = 0.75)
text(x = 2000, y = 8, bquote(.(release_yr_coef)~kg~ha^-1~yr^-1))
mtext("Adjusting for yield erosion", adj = 0)

polygon(x = c(yrs, rev(yrs)),
        y = c(ints2[1,],rev(ints2[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

@
% Change from a./b. to accounting for erosion, More details about how
% yield was corrected - 
\caption{The relationship between yield and the release year of medium
grain rice cultivars for the last 6 years
(\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated by 
correcting for yield erosion over time. Yield erosion was corrected
for by adjusting each observed yield by estimated amount lost 
  (\Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} after release).}
\label{fig:w_erosion}
\end{figure}
\end{comment}

\section{Discussion}

In California, the cultivars with broad adaptation and adoption over
the 30+ year period of this study are M-202, M-206, and M-105. Cold
tolerance is a major problem for much of the California rice producing
region ~\citep{espe2017}, although it is a bigger problem in the
southern portion of the Sacramento Valley due to the influence of the
ocean winds through the Delta ~\citep{espe16a}. These three
cultivars have a higher degree of cold tolerance, high yield
potential, and relatively short duration and are thus broadly grown
with over 70\% of the area in one of these cultivars during the study
period ~citep{ucce}. Importantly, not all medium grain cultivars that were
released were purported to have higher yields than previous
cultivars. Some cultivars had special traits, but upon release had
similar yields to more broadly adapted cultivars. Examples of this
include M-208, which had resistance to blast disease (scientific name)
and M-402 which is a premium quality variety. Other varieties such as
M-205 and M-209, while high yielding, do not have the cold tolerance
of the other cultivars and are thus only recommended for the northern
portion of the Sacramento Valley.

Looking at all the cultivars that have been released, there is no
conclusive evidence of increased genetic yield potential in medium
grain rice cultivars released by CCRRF's breeding improvement program
over the period \Sexpr{min(mod_data[,'year'])} to
\Sexpr{max(mod_data[,'year'])}, which has stayed relatively constant
at \Sexpr{round(mean(mod_data[,"yield_kg"]), 2)} Mg
ha\tsups{-1}. Since the estimated yield improvement spans from
\Sexpr{rel_CI[1]} to \Sexpr{rel_CI[2]} kg ha\tsups{-1} year\tsups{-1}
(Fig. \ref{fig:RES_year}; Table \ref{tbl:coef}),
both increases and decreases in the genetic yield potential over the
study period are plausible. The estimated yield
improvement of \Sexpr{only_release_coef} kg ha\tsups{-1}
year\tsups{-1} (Fig. \ref{fig:wo_erosion}) that does not account for
yield erosion is lower than the 37 to 60 kg ha\tsups{-1}
year\tsups{-1} estimated by~\citet{Omar2014} in side-by-side
comparisons. Rather, this analysis suggests that the CCRRF breeding
program has maintained but not substantially improved the yield
potential in the face of changing biotic and abiotic pressures. This
conclusion holds true even when evaluating the cultivars which have
been breed for broad adaption and high yield potential (M-202, M-206
and M-105). Yields of these varieties at time of release have all
yielded between
\Sexpr{tmp=mod_data[mod_data[,"yrs_in_trial"]<=0&mod_data[,"id"]%in%c("M202","M206","M105"),];round(min(by(tmp,tmp[,"id"],function(x)mean(x[,"yield_kg"]))/1000),2)}
and
\Sexpr{round(min(by(tmp,tmp[,"id"],function(x)mean(x[,"yield_kg"]))/1000),2)}
Mg ha\tsups{-1} with no indication of yields increasing
(Fig. \ref{fig:first_yrs}). While these yields have not increased,
M-205 did have higher yield potential; however, this variety is not
broadly adapted due to poor cold tolerance.

Yields of a particular cultivar may decline overtime for a number of
reasons. First, ~\citet{Nie2009} reported yield declines in rice due
to continuously flooded conditions over a long time period which could
be overcome with improved N management. This is likely not the case
here as the fields used in this study grow rice only once season a
year and tend to be non-flooded for the remainder of the
year. Furthermore, at this location and unlike most CA rice fields,
the fields are left fallow every other year. Second, changing climate
may influence yields with either warmer~\citep{van2013} or
cooler (Espe et al., 2017) temperatures reducing yields. Again, this
is unlikely to be the case in this situation as Espe et al. (2017)
found no significant changes in growing season temperatures (either
high, low or average temps) the Sacramento Valley during the study
period.  Thirdly, yield erosion of a cultivar as has been reported for
the cultivar IR8 at IRRI~\citep{Peng1998, Peng2000, Peng2010}. Biotic
and abiotic forces have previously been speculated to be responsible
for yield erosion, since the cultivar's genetics remain stable from
the time of selection but the conditions the cultivar may experience
over time are continuously changing~\citep{Peng1998, Peng2000,
Peng2010}. We estimate that yield erosion has decreased the yield of
medium grain rice cultivars released by the CCRRF over the last 30+
years by an average of \Sexpr{yrs_coef} Mg ha\tsups{-1}, potentially
offsetting any gains made by the crop improvement program. However, it
is not clear what is responsible for this erosion. It is unlikely
biotic factors as pest and disease pressure are low in this
Mediterranean climate. Furthermore, these fields are fallowed every
other year and the residues are burned, further lowering the potential
for pest and disease pressure. While efforts should be made to
identify the cause of erosion in this environment, this observation
highlights both the need for maintenance of plant breeding and the
difficulty of increasing genetic yield potential, especially in
systems such as CA where yield is approaching the physiological limit
\citep{espe16a}. Without maintenance breeding (i.e., continuously
selecting cultivars well adapted to current conditions), CA rice
production systems are predicted to have lost an estimated
\Sexpr{round(yrs_coef*(-36/1000),1)} Mg ha\tsups{-1} grain yield over
the study period.

A goal of many breeding programs is to not only maintain the benchmark
yield of the previous ``standard'' but to surpass it, especially if we
are to meet the target of 50\% increased food production by 2050
\citep{Foley2011a}. Given the estimated yield erosion, to hit an
ambitious target of 0.5 Mg ha\tsups{-1} (approximately 5\%) increase
in yield per release (with roughly 4 years between releases), a new
cultivar needs to yield an additional \Sexpr{round(yrs_coef*-4,-1)} kg
ha\tsups{-1} in side-by-side comparison in addition to the 0.5 Mg
ha\tsups{-1} target improvement. Often side-by-side yield improvements
show an improvement in grain yield of 2\% per release, in which case
no absolute improvements in yield have been made beyond maintaining
yield. The evidence here is consistent with the observation that the
genetic yield potential has not increased substantially in the last
30+ years (Fig. \ref{fig:first_yrs}).

\begin{figure}[H]
<<fig = TRUE, echo = FALSE, width=5, height=5>>=

plot(1, type = "n",
     # xaxt = "n",
     xlab = "Years",
     ylab = expression("Grain yield ( Mg ha"^-1~")"),
     ylim = c(8.5,11.5),
     xlim = c(0,40),
     xaxs = "i")

# arrows(12, 8.2, 27, 8.2, xpd = TRUE, lwd = 2)
abline(a = 10, b = fixef(RES_full)[3], lty = 2, lwd = 2)
text(25, 10.85, "(3) target above maintenance", cex = 0.8,
     srt = 18)#((atan(fixef(RES_full)[3] + 0.06) * 180))/pi)
abline(h = 10, lty = 3, lwd = 2)
text(25, 10.1, "(1) current situation", cex = 0.8)
abline(a = 10, b = fixef(RES_full)[3] + 0.06, lty = 4, lwd = 2)
text(25, 9.35, "(2) no maintenance breeding", cex = 0.8,
     srt = -18)

@ 
\caption{A conceptual diagram highlighting three scenarios: (1)
maintenance breeding, where losses over time are offset by
improvements via plant improvement, (2) no maintenance breeding, where
grain yield slowly erodes over time (as estimated from this study),
and (3) absolute improvement in grain yield (based on ``typical''
yield goals of 3\% increase per release), where gains in grain yield
exceed the losses over time.}
\label{fig:theory}
\end{figure}

There are several cultivars that were found to have better than
average yields, even after correcting for yield erosion. However,
these cultivars (M-201, M-204, M-205, and possibly M-209;
Fig. \ref{fig:ids_comp})) are not widely adopted. It is known that
cool temperatures during the booting stage can have a large impact on
grain yields \citep{espe2017}, and the RES
experiences warmer temperatures than many other rice growing areas in
CA \citep{espe2017}. These cultivars may perform well at the RES but
not widely in CA. Therefore, it should be noted that while our
investigation has been concerned about increasing yield potential, the
estimates here might not capture yield stability. Yield stability is
notoriously difficult to quantify \citep[e.g.,][]{lobell2011}, hence
further investigation is needed to explore how yield stability has
changed over the study period.

To be clear, our investigation here is narrowly focused on the
rough-rice/field grain yield. Other advancements due to plant
improvement, including increased quality \citep{mckenzie1993} and
milling yield \citep{mckenzie1994}, have been made over the study
period and are important. Also, crop duration has been reduced. For
example, during this study period, M-202 and M-206 have been the
predominate cultivars grown in CA, and while their yield potentials
were similar at the time of release (Fig. 5), M-206 matures about 6
days earlier than M-202 (Table \ref{tbl:ids}). Shorter duration is very important
in CA due to an already short growing season and water limitations
(shorter duration cultivars would presumably require less irrigation
water).

California state average yields have increased slowly
despite there being little to no gains in the breeding trials
(Fig. \ref{fig:state_yield}). The evidence here suggests mechanisms
other than increased genetic yield potential are responsible for these gains
(Fig. \ref{fig:compare_yrs}). Improvement in agronomic management,
including more efficient nutrient management \citep{Lundy2012,
Lundy2015, linquist2009}, weed management \citep{Saito2008,
Pittelkow2012, linquist2008}, and other factors may all be contributing to
increases in the state average yield. However, these gains are far less
than those experienced during the preceding era, and by their
nature agronomic gains cannot surpass the genetic yield potential
\citep{Fischer2015}. Judging from yields achieved at the RES
(Figs. \ref{fig:RES_year} \& \ref{fig:compare_yrs}), a
researcher-managed production system implementing best management
practices in more controlled situations compared to farmers' fields, state
average yields (approximately 10 Mg ha\tsups{-1} at the time of
writing; Fig. \ref{fig:state_yield}) may be approaching the current
genetic limit. Assuming that the level of management possible in small
experimental plots cannot be attained in farmers' fields and hence
experimental yields represent the attainable yield ceiling
\citep{Fischer2015}, agronomic gains in farmer's fields are nearing
their potential, consistent with physiological model-based estimates
of the attainable yield gap in CA rice production systems
\citep{espe16a}.

Planning to meet future
demand necessitates increases in grain yield \citep{Godfray2010,
Foley2011a, Mueller2012}. Meeting the challenge to produce more grain
without expansion of land under cultivation will rely on new genetics
with increased yield potential. The evidence here suggests that to
accomplish true increases over time, the improvement target needs take
into account erosion in the current ``standard'' cultivars' yield
over time (Fig. \ref{fig:theory}). While side-by-side comparisons will
continue to be useful, comparisons across years to each cultivar's
benchmark yield at the time of release should provide information on
the absolute improvements being made in yield potential over time. New
technologies may need to be adapted to further increase yield
potential, including hybrid rice.

\section{Conclusions}

Here we show that side-by-side cultivar comparisons do not necessarily
indicate gains in yield potential as they do not account for yield
erosion. Quantifying changes in yield potential requires also
examining historical benchmarks. Improvement in grain yield is
essential to meeting the challenges of feeding a growing population in
a changing world. To achieve these improvements, we discuss here that
(1) continuous maintenance via plant improvement is required, without
which yields would decrease over time, (2) the challenge of improving
genetic yield potential is substantial, since absolute gains require
improvement in yield in excess of the maintenance threshold, and (3)
increasing yield through improved management (agronomy) cannot exceed
the ceiling of genetic yield potential. Taken together, these
conclusions underscore both the necessity and substantial challenges
facing rice breeding. We suggest here that historical benchmarks be
used in to evaluate promising rice cultivars in addition to
side-by-side comparisons to maintain perspective of the absolute yield
gains being made.

\newpage

\section{Tables}

<<echo=FALSE, results=tex>>=

CIs = posterior_interval(RES_full)[1:3,] * scl
CIs[1,] = round((CIs[1,] + center)/1000, 1)
CIs = round(CIs, 1)
med_eff = fixef(RES_full) * scl
med_eff[1] = round((med_eff[1] + center)/1000, 1)
med_eff = round(med_eff, 1)

RES_coef = cbind(med_eff, paste(CIs[,1],CIs[,2], sep = " to "))
rownames(RES_coef) = c("Intercept (Mg ha\\tsups{-1})",
                       "Yield improvement (kg ha\\tsups{-1} year\\tsups{-1})",
                       "Yield erosion (kg ha\\tsups{-1} year\\tsups{-1})")
colnames(RES_coef) = c("Median", "90\\% Credible interval")

print(xtable(RES_coef,
             caption = "The estimated median fixed effects and 90\\% credible intervals for the improvement in yield of medium grain rice cultivars developed by the California Cooperative Rice Research Foundation, California, USA over the time period 1984 -- 2016.",
             label = "tbl:coef",
             align = "lcc"),
      caption.placement="top",
      booktabs=TRUE,
      sanitize.colnames.function = function(x){x},
      sanitize.rownames.function = function(x){x})
@
\newpage

<<echo=FALSE, results=tex>>=
ids = aggregate(release_yr ~ id, data = mod_data, unique)
ids$dth = as.integer(round(aggregate(dth ~ id, data = mod_data, mean, na.rm = TRUE)$dth, 0))
colnames(ids) = c("Cultivar", "Release Year", "Days to 50% Heading")
ids = ids[order(ids[,2]),]
ids[,2] = as.integer(ids[,2])
ids[,1] = gsub("M", "M-", ids[,1])

print(xtable(ids,
             caption = "All named medium grain rice cultivars and their associated release year planted at the Rice Experiment Station (RES), Biggs, California from 1984 -- 2015.",
             label = "tbl:ids"),
      include.rownames = FALSE,
      booktabs = TRUE,
      caption.placement = "top")

@

% Notes: Duration has gotten shorter. Etc. Rearrange - End
% introduction with evidence of erosion, quatify the effect on yield
% increases.

\newpage

\section{References}\label{references}

\setlength{\parindent}{-0.2in} \setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}

\bibliography{breeding}

\newpage

\section{Supplemental tables and
figures}\label{supplemental-tables-and-figures}

\beginsupplement

\begin{figure}[H]

<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=
tars <- c("M202","M206")

plot_data = lapply(tars, function(x) {
    dd = mod_data[mod_data$id == x & mod_data$site == "RES",]
    tmp = prep_plot_data(dd$yield_kg / 1000, dd$yrs_in_trial)
    tmp$yrs_in_trial = as.numeric(row.names(tmp))
    tmp
    })

plot(mean ~ yrs_in_trial,
     data = plot_data[[1]],
     xlim = c(-6, 32), ylim = c(6.5,14.5),
     pch = 16, #main = gsub("M","M-", tars[i]),
     ylab = expression("Yield ( kg ha"^-1~")"),
     xlab = "Years after release")
segments(x0 = plot_data[[1]]$yrs_in_trial,
         y0 = plot_data[[1]]$mean + plot_data[[1]]$se,
         y1 = plot_data[[1]]$mean - plot_data[[1]]$se)
points(mean ~ yrs_in_trial,
       data = plot_data[[2]],
       pch = 17, col = "grey40")
segments(x0 = plot_data[[2]]$yrs_in_trial,
         y0 = plot_data[[2]]$mean + plot_data[[2]]$se,
         y1 = plot_data[[2]]$mean - plot_data[[2]]$se,
         col = "grey40")
abline(lm(yield_kg/1000 ~ yrs_in_trial,
          data = mod_data[(mod_data$id %in% tars & mod_data$site == "RES"),]),
       lty = 1, col = "grey50")
legend("topright", legend = gsub("M","M-", tars),
       pch = c(16, 17), col = c("black","grey40"))

@
\caption{The performance of two major cultivars at the CA Rice
  Experiment Station (RES) since the year each cultivar was
  released. Both M-202 ($\bullet$) and M-206 ($\blacktriangle$) have similar
  intercepts, suggesting similar yield at the time of release and an
  apparent yield advantage of the newer cultivar, M-206, might be
  explained by a yield loss of M-202 over time. Points represent the
  mean, and bars (|) the standard error.}
\label{fig:compare_since_release}
\end{figure}

\newpage

\begin{figure}[H]
<<fig=TRUE, width = 8, height = 5, out.width = '85%', echo=FALSE>>=
tars <- c("M202","M206")
par(mfrow = c(1, length(tars)),
    mar = c(4,0,2,0),
    oma = c(4,4,0.5,0.5))

plot_data = lapply(tars, function(x) {
    dd = mod_data[mod_data$id == x & mod_data$site == "RES",]
    tmp = prep_plot_data(dd$yield_kg/1000 , dd$year)
    tmp$year = as.numeric(row.names(tmp))
    tmp
    })

invisible(sapply(1:2, function(i){
    plot(mean ~ year,
         data = plot_data[[i]],
         xlim = c(1985, 2016), ylim = c(7.5,13.5),
         pch = 16, main = gsub("M","M-", tars[i]),
         ylab = "",
         yaxt = "n",
         xlab = "Year",
         cex.axis = 0.8)
    segments(x0 = plot_data[[i]]$year,
             y0 = plot_data[[i]]$mean + plot_data[[i]]$se,
             y1 = plot_data[[i]]$mean - plot_data[[i]]$se)
    if(i == 1){
        axis(2, cex.axis = 0.8,)
        mtext(expression("Yield ( Mg ha"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }
    # abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000 -
    #            (2000 * yrs_coef/1000),
    #        b = yrs_coef/1000,
    #        lty = 1)
    lines(as.numeric(rownames(plot_data[[i]])),
          11.2 + (1:nrow(plot_data[[i]])) * (yrs_coef/1000))
    # abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000,
           # b = yrs_coef/1000)

    # abline(h = mean(mod_data$yield_kg[(mod_data$id == tars[i] &
    #                                    mod_data$site == "RES")] ),
    #        lty = 1, col = "grey50")
    # mod = lm(mean ~ year, data = plot_data[[i]])
    # tmp = predict(mod)
    # if(i == 1){
    #     abline(mod)
    # }else{
    #     lines(as.numeric(names(tmp)), tmp)
    # }
    abline(h = 11.2, lty = 2)
    # text(2000, 7.5, bquote(.(round(coef(mod)[2] * 1000, 1))~kg~ha^-1~year^-1))
}))

@
\caption{The performance of two major cultivars over the period
\Sexpr{min(mod_data[,'year'])} -- \Sexpr{max(mod_data[,'year'])}. The
solid line is the estimated yield erosion (\Sexpr{yrs_coef} kg
ha\tsups{-1} year\tsups{-1}), while the dashed line is the published
3-year average yield of M-202 at its release (1982-1984). Both
cultivars have a similar yield at their respective year of release,
and similar decreases over time.}
\label{fig:compare_yrs}
\end{figure}
\newpage
% \begin{figure}[H]
% <<fig=TRUE, echo=FALSE>>=
% tars <- c("M202","M206")
% comp_yrs = 2004:2005

% boxplot(yield_kg/1000 ~ id, data = mod_data[(mod_data$id %in% tars) &
%                                             (mod_data$year %in% comp_yrs),],
%         ylab = expression("Grain yield ( Mg ha"^-1~")"))

% @
% \caption{An example of a typical side-by-side comparison between a older (M-202)
%   and newer (M-206) cultivar. Shown is the performance of these cultivars at the CA Rice
%   Experiment Station (RES) over the period \Sexpr{min(comp_yrs)} --
%   \Sexpr{max(comp_yrs)}. The newer
%   cultivar, M-206, appears to have a yield advantage over the older
%   cultivar, M-202, though this comparison does not account for the 10
%   year difference in years since release between the cultivars.}
% \label{fig:compare_vars}
% \end{figure}
% \newpage

\begin{figure}[H]
<<echo = FALSE, fig = TRUE, width = 7, height = 6>>=
library(ggplot2)
library(ggthemes)
ggplot(mod_data, aes(x = year, y = yield_kg/1000)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_smooth(method = "lm", color = "black") +
    theme_few() +
    xlab("Year") +
    ylab(expression("Grain yield ( Mg ha"^-1~")")) + 
    facet_wrap(~id, ncol = 4)

@
\caption{The yield of all medium grain released cultivars at the
  California Cooperative Rice Research Foundation Rice Experiment
  Station over the period \Sexpr{min(mod_data[,'year'])} --
  \Sexpr{max(mod_data[,'year'])}.}
\label{fig:all_cult}
\end{figure}

\newpage
\begin{figure}[ht]
<<echo=FALSE, fig = TRUE>>=

dotchart(id_est[,1],
         labels = gsub("M", "M-", row.names(id_est)),
         pch = 16,
         xlim = c(min(id_est[,2]),max(id_est[,3])),
         xlab = expression("Estimated difference from average ( Mg ha"^-1~")"))
segments(x0 = id_est[,2], x1 = id_est[,3], y0 = 1:nrow(id_est))
abline(v = 0, lty = 2)
@
\caption{The marginal estimates for the difference between cultivars,
  averaged over the uncertainty in the fixed effect parameters (i.e.,
  yield erosion and annual yield gains).}
\label{fig:ids_comp}
\end{figure}

\newpage

\end{document}