### 2017-11-13 output/breed_paper.Rnw
Final edits.

\documentclass[12pt,]{elsarticle}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{longtable,booktabs}
\usepackage{comment}
\usepackage{graphicx,grffile, textcomp}
\usepackage{setspace}
\usepackage{etoolbox}
\usepackage{float}
\doublespacing
\usepackage{lineno}
\usepackage{lscape} %% Harvard
\bibliographystyle{model2-names}\biboptions{authoryear}
\pagenumbering{gobble}
%% For figure numbers in supplementary info
%% from http://bytesizebio.net/2013/03/11/adding-supplementary-tables-and-figures-in-latex/
\newcommand{\beginsupplement}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
}

\newcommand{\tsups}{\textsuperscript}
\usepackage{Sweave}
\SweaveOpts{prefix.string=graphics/report, png=TRUE}

\begin{document}

\begin{frontmatter}

\title{Rice yield improvements through plant breeding are offset by
  inherent yield declines over time}
\author[ucd]{Matthew B. Espe\corref{corresauthor}}
\cortext[corresauthor]{Corresponding author} \ead{mespe@ucdavis.edu}
\author[ucd]{Jim E. Hill}
\author[ucce2]{Michelle Leinfelder-Miles}
\author[ucce]{Luis A. Espino}
\author[ucce]{Randall Mutters}
\author[ucd]{David Mackill}
\author[ucd]{Chris van Kessel}
\author[ucd]{Bruce A. Linquist}

\address[ucd]{Dept. of Plant Sciences, University of California --
Davis, Davis CA 95616, USA}

\address[ucce]{University of California Cooperative Extension,
Oroville, CA 95965, USA}

\address[ucce]{University of California Cooperative Extension,
Colusa, CA 95932, USA}

\address[ucce2]{University of California Cooperative Extension,
Stockton, CA 95206, USA}

\begin{abstract}

Meeting the challenge of feeding a growing population with limited
resources will require increasing the yield potential of staple crops,
such as rice. Yet many high-yielding, intensive production systems
have experienced slow rates of yield improvement in recent years
despite a demonstrated increase in the yield potential of new crop
cultivars. We analyzed experimental data from one such cropping
system, i.e., California (CA) rice, in order to quantify improvements
made in the genetic yield potential obtained through plant
breeding. California rice systems are among the highest in the world
and close to maximum yield potential. Specifically, the hypothesis was
tested that if rice cultivar yields decline over time (termed yield
erosion) then apparent yield increases in side-by-side yield
comparison tests will not reflect increases in yield potential. This
hypothesis was tested using 33 years of experimental yield data from
the California Cooperative Rice Research Foundation Rice Experiment
Station. Based on side-by-side comparisons of old and new rice
cultivars which do not consider yield erosion, there was an apparent
increase in yield. However, the yields of older cultivars eroded at an
estimated rate of 29.3 kg ha\tsups{-1} year\tsups{-1} (90\% credible
interval -4.4 to -53.3) after initial selection. Once this effect was
considered, the yield advantage of newer cultivars over old was
uncertain (-3.3 kg ha\tsups{-1} year\tsups{-1}, 90\% credible interval
-36.1 to 31.5). These results highlight 1) the importance of
continuous crop improvement and deployment of new cultivars simply to
maintain existing yields, and 2) to increase the genetic yield
potential, higher yield targets are needed. Importantly, when breeding
near the yield potential, despite the limited yield gains, significant
advances in improving quality and reducing crop duration have been
made.

\end{abstract}

\begin{keyword} rice, yield potential, breeding, yield erosion
\end{keyword}

\end{frontmatter}

\newpage

\textbf{Abbreviations}

CA: California

CI: Credible Interval

RES: Rice Experiment Station

CCRRF: California Cooperative Rice Research Foundation

MC: Moisture Content

\newpage

\linenumbers

<<echo=FALSE>>=
library(rstan)
library(rstanarm)
library(lme4)
library(xtable)

load("../data/model_data.rda")
load("h1_fit.Rda")
load("bivar_fit.Rda")
load("../data/all_vt_weather.Rda")

prep_plot_data = function(x, idx)
{
    as.data.frame(do.call(rbind, tapply(x, idx, function(x){
        c(mean = mean(x), se = (sd(x)/sqrt(length(x))))
    })))
}

release_yr_coef = round(fixef(RES_full)[2] * scl, 1)
yrs_coef = round(fixef(RES_full)[3] * scl, 1)
only_release_coef = round(fixef(RES_rel_fit)[2]*scl, 1)

RES_CI <- round(posterior_interval(RES_full, pars = "yrs_in_trial") * scl, 1)
rel_CI = round(posterior_interval(RES_full, pars = "release_yr_c") * scl, 1)
only_release_CI = round(posterior_interval(RES_rel_fit, pars = "release_yr_c") * scl, 1)

m202_v_206 = round(yrs_coef * -19, -1)

id_est = ranef(RES_full)$id
id_est = cbind(id_est, posterior_interval(RES_full, regex_pars = "id:M"))
idx = order(id_est[,1], decreasing = FALSE)
id_est = id_est[idx,]

@

\section{Introduction}\label{intro}

\setlength{\parindent}{1cm}

Constraints on arable land are increasing simultaneous with the need
to increase total food production to meet a growing demand
\citep{Foley2011a, Godfray2010, Mueller2012}, which has necessitated
harvesting more grain per unit land area~\citep{Lobell2009,
Tittonell2014}. Historically, agricultural research has been
successful in staving off the ``Malthusian catastrophe'' of demand
surpassing supply via continued yield improvement of staple
crops. However, many production systems are experiencing plateaus in
grain yield~\citep{Grassini2013}. If this trend continues,
improvements per unit land area are no longer possible, and an
increase in the area under cultivation will be needed to meet food
demand, which carries with it undesirable ecological
implications~\citep{Foley2011a, tilman1999}. Therefore, it is of
critical importance to better understand why the rate of increase in
grain yields has declined or leveled off in intensified production
systems.

California (CA) rice represents one such production system. Rice is
grown primarily in the Sacramento Valley, which is characterized by
having a Mediterranean climate with long days, a dry growing season
relatively free of pests and diseases. These conditions lead to some
of the highest yields in the world~\citep{FAO2016}. Production in CA
is predominately focused on premium quality medium grain
\textit{japonica} cultivars (e.g., CalRose rice), and rice from CA is
recognized globally for its quality
(\url{http://agfax.com/2015/11/03/rice-calrose-wins-best-rice-world-competition/}). Most
cultivars in use in CA are developed by the California Cooperative
Rice Research Foundation (CCRRF; a collaboration between the
University of California, the USDA-Agricultural Research Service, and
farmer-funded research). In part due to improvements in rice genetics,
rice yields in CA increased rapidly during the period from 1920 --
1990; however, since the 1990s, the rate of yield increase has slowed
(Fig. \ref{fig:state_yield}) despite continuous crop
improvement. Based on a yield-gap analysis, on-farm yields in the
major rice growing region of CA is 73 to 76\% of maximum yield
potential~\citep{espe16a}. In highly intensive systems such as this,
\citet{grassini2011} have shown that farmers are capable of attaining
85\% of the maximum yield potential. Thus, given that farmers are near
the attainable yield potential, increases in yield are likely to be
slow and \citet{espe16a} reported that indeed this was the case
with yields on average increasing by about 50 to 62 kg ha\tsups{-1}
year\tsups{-1} between 1999 and 2014.

Certainly one question in looking at this situation is, ``do new
cultivars have increased yield potential''?  Like many breeding
programs, one of the goals of the CCRRF is to increase yields through
breeding cultivars with increased yield potential. However, when the
popular cultivar M-202 was released in 1985, it was reported to yield
an average of 11 Mg ha\tsups{-1}~\citep[14\% moisture content
(MC);][]{m202}. In 2015, the reported average yield of the newest
released cultivar M-209 was 10.8 Mg ha\tsups{-1}~\citep[14\%
MC;][]{m209}. However, in annual breeding reports and cultivar release
announcements, data based on yield performance in side-by-side cultivar
trials show that new cultivars usually out yield the older
cultivars. In other rice systems, a decrease in a cultivar's yield
performance over time has been observed~\citep{Peng1998, Peng2010,
Peng2000, Datta1995}, often termed yield erosion~\citep{Peng1998}. It
has been speculated that this yield erosion is due to the inability of
a cultivar to adapt to changing biotic and abiotic
conditions~\citep{Peng1998, Peng2010, Peng2000, Datta1995}. Yield
erosion may explain the apparent contradiction above, in which
side-by-side yield comparisons show yield improvements in cultivars
while overall there may be little change in yield potential. To
address this issue we tested the hypothesis that if rice cultivar
yields erode over time then apparent yield increases in side-by-side
yield comparison tests will not reflect the true magnitude of yield
changes over time.

\begin{figure}[H]
<<fig=TRUE, echo=FALSE>>=
CA_data = read.csv("../data/CA_rice_yield.csv", stringsAsFactors = FALSE)
CA_data$Value = as.numeric(gsub(",", "", CA_data$Value))/1000
CA_data = CA_data[order(CA_data$Year),]
plot(Value * 1.12 ~ Year, CA_data,
     # type = "l",
     pch = 16,
     xlab = "Year",
     ylab = expression("State avg. grain yield ( Mg ha"^-1~")"))
# mv_avg = filter(c(rep(mean(CA_data$Value[1:10] * 1.12), 5),
#                   CA_data$Value * 1.12,
#                   rep(mean(CA_data$Value[(nrow(CA_data) - 10):nrow(CA_data)] * 1.12), 5)),
#                 rep(1/11,11))
# lines(x = c(1907:1911, CA_data$Year, 2017:2021), y = mv_avg)

@
\caption{Average rice yields over the period \Sexpr{min(CA_data$Year)}
to \Sexpr{max(CA_data$Year)} for California, USA (US Dept. of
Agriculture National Agricultural Statistics Service, 2017).}
\label{fig:state_yield}
\end{figure}

\section{Methods}\label{methods}

\subsection{Site and data description}

The CCRRF Rice Experiment Station (RES) is located near Biggs, CA
(39.4648, -121.7342), and has been the central location for CCRRF's
efforts to develop improved rice cultivars adapted to CA since
1969. The California Statewide Variety Trails evaluate current and
promising cultivars at the RES and six to eight on-farm trials around
the state each year. To avoid potential complications due to site,
climate, and management differences at the farmer-managed trials
between years, we focused our investigation on the RES. The RES is
researcher-managed and plants the majority of the experimental plots
for the statewide program (approximately 3 to 4 times the number of
experimental plots compared to other on-farm locations). The climate
at the RES is Mediterranean, characterized by mild winters during
which most of the annual precipitation occurs (\Sexpr{18*24.5} to
\Sexpr{25*24.5} mm annually) and warm summers largely free of
precipitation events. Soils at the RES are classified as
Esquon-Neerdobe clays with roughly 2\% soil organic matter in the top
15cm~\citep{ssurgo}.

As part of the annual Statewide Variety Trials, newly developed
entries are tested against officially released cultivars (checks) at
the RES in several trials spanning multiple planting dates per
year. For the purposes of this study, we concentrated on released
medium grain cultivars, as these cultivars constitute approximately
90\% of the planted rice area in CA. Newly developed entries are
typically evaluated in these trials for 3 to 5 years prior to their
official release. For these cultivars planted in breeding trials at
the RES from \Sexpr{min(mod_data[,'year'])} --
\Sexpr{max(mod_data[,'year'])} plot level observations (replicated 3
to 4 times in a CRB design) were collected leading to a final data set
of \Sexpr{format(nrow(mod_data), big.mark = ',')} observations
representing \Sexpr{length(unique(mod_data[,'id']))} cultivars and
included every publicly released medium grain cultivar developed by
the CCRRF from \Sexpr{min(mod_data[,'release_yr'])} --
\Sexpr{max(mod_data[,'release_yr'])}. For each cultivar, the official
year of public release and the time elapsed (in years) since release
was determined.

Experimental plots at the RES are managed similar to University of
California Cooperative Extension prescribed best management practice
for land preparation, fertility rates, and pest and disease
control~\citep{ucce}. As is typical for water-seeded systems, the
plots were planted by pre-germinating rice seed prior to
direct-seeding into pre-flooded fields. Plots were either 14.0 or 18.5
m\tsups{2}, and were harvested after physiological maturity using a
small-plot combine. All yields were converted to 14\% MC prior to
reporting.

\subsection{Statistical analysis}

To test the hypothesis, the influence of breeding and yield erosion
were quantified on yield over time using three nested Bayesian hierarchical
models. First, yield was modeled as a response to the number of years since
the cultivar was officially released and the year of release using a
mixed-effects linear regression. To account for similarities between
years and between cultivars, year and cultivar were included in the
model as random effects.

\begin{equation}
\begin{split}
 yield_i = Intercept + (Rate_{yield\ erosion\ i}* {years\ since\
   release\ i}) \\+ (Rate_{yield\ improvement\ i}* {release\ year\ i}) 
 + \beta_{year\ i} + \beta_{cultivar\ i} \\
\end{split}
\end{equation}

Where $yield_i$ is the grain yield for plot $i$, the Intercept is the
average yield for the RES across cultivars and years, $Rate_{yield\
erosion\ i}$ is rate of yield erosion per ${years\ since\ release\ i}$
(negative for years prior to release, zero at the year of release),
$Rate_{yield\ improvement\ i}$ is the rate of yield improvement per
year (${release\ year\ i}$), $\beta_{year}$ is the random effect of
year, and $\beta_{cultivar}$ is the random effect of cultivar. This
full model was modified to further explore yield dynamics over
time. For comparison to estimates where yield erosion is not taken
into account, the model was used as described above except omitting
the $Rate_{yield\ erosion\ i}$ term (Model 2). Lastly, a model was fit
to estimate if there were differences between cultivars in the rate of
yield erosion by adding a random slope by cultivar (Model 3).

Data were processed and models fit in R, an environment for
statistical computing~\citep{R}. Models were fit using the `rstanarm'
package~\citep{rstanarm}, an interface to Stan, a language for
probabilistic programming~\citep{stan}. All model diagnostics,
including $\hat{R}$, effective size, and posterior predictive checks
were examined before reporting results. To test goodness of fit, the
full model was compared to models containing only the years since
release or the year of release using the `loo' package~\citep{loo}, an
efficient means of conducting leave-one-out cross validation
\citep{loo_paper}. The 90\% credible interval (90\% CI), defined as
the interval containing 90\% of the distribution of estimated credible
parameter values given the data, was calculated as a measure of
uncertainty in the parameter estimates using the quantile method. The
90\% CI is more stable to sample to sample variance
and hence is preferable to the 95\% interval~\citep{rstanarm}. To
verify the estimates, the analysis was also conducted using classical
methods, specifically frequentist mixed-effects models fit by maximum
likelihood (see supplemental material for details and model
results). Complete data and code used for this analysis is available
at \url{github.com/mespe/rice_advances.git}.

\section{Results}\label{results}

\begin{figure}[H]
<<fig=TRUE, width = 6, height = 5, out.width = '85%', echo=FALSE>>=

idx = mod_data[,"yrs_in_trial"] %in% -2:0

plot_data = prep_plot_data(mod_data[idx,"yield_kg"] / 1000 ,
                           mod_data[idx, "id"])

plot_data[,"yr"] = tapply(mod_data[idx, "release_yr"],
                          mod_data[idx,"id"], unique)

plot(mean ~ yr, data = plot_data,
     pch = 16,
     xlab = "Year of release",
     ylab = expression("Yield ( Mg ha"^-1~")"),
     ylim = c(9,13),
     xlim = c(1984, 2018.5))

segments(x0 = plot_data[,"yr"],
         y0 = plot_data[,"mean"] + plot_data[,"se"],
         y1 = plot_data[,"mean"] - plot_data[,"se"])

#Used locator() it interactively find label points 
aa = readRDS("f2_text_pos.rds")
text(x = aa$x, y = aa$y,
     labels = gsub("M", "M-", row.names(plot_data)),
     pos = 4, cex = 0.75)
#### Polygons
yrs <- 1975:2025

pp <- posterior_linpred(RES_full, re.form = NA,
                        newdata = data.frame(release_yr_c = yrs - 2000,
                                             yrs_in_trial = 0),
                        prob = 0.5)
pp <-  (pp * scl + center) / 1000
ints2 = apply(pp, 2, quantile, p = c(0.05, 0.95))

abline(a = ((((fixef(RES_full)[1] * scl) + center) ) -
           (2000 * release_yr_coef)) / 1000,
       b = release_yr_coef/1000,
       lty = 2)
polygon(x = c(yrs, rev(yrs)),
        y = c(ints2[1,],rev(ints2[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

@

\caption{The experimental yield of medium grain rice cultivars for the
  two years prior to each cultivar's public release. Points
($\bullet$) and bars (|) represent the mean and standard error of the
raw data, respectively. The dashed line and shaded area are the
genetic yield improvement and credible interval as estimated by the model.}
\label{fig:first_yrs}
\end{figure}

\subsection{Cultivar yields at time of release}

There was general agreement between the estimates from the Bayesian
and classical statistical approaches for all models, with the
estimates being similar between similar models. The Bayesian estimates
are presented here due to the inherent flexibility of Bayesian
inference, though the classical estimates can be found in the
supplemental material. Comparing the cultivars' yield during each's
initial time (average 3 years) in the experimental trials just before
being released showed that across cultivars, yields ranged from
\Sexpr{round(plot_data["M207", "mean"], 2)} (M-207) to
\Sexpr{round(plot_data["M205", "mean"], 2)} Mg ha\tsups{-1} (M-205)
(Fig \ref{fig:first_yrs}. Comparing the yields from only the first few
years for each cultivar over the study period does not provide
evidence of genetic yield improvements during the 33-year study
period. Individual exceptions to this were two cultivars (M-204 and
M-205) that had higher yields than the others.

\subsection{Yield erosion}
\begin{figure}[H]

<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=
yrs_in <- -10:40
pp <- posterior_linpred(RES_full, re.form = NA,
                        newdata = data.frame(yrs_in_trial = yrs_in,
                                             release_yr_c = 0))
pp <-  (pp * scl + center) / 1000 
ints = apply(pp, 2, quantile, p = c(0.05, 0.95))

plot_data = prep_plot_data(mod_data$yield_kg / 1000,
                           mod_data$yrs_in_trial)

plot_data$tmp = as.numeric(row.names(plot_data))
colnames(plot_data) = c("yield_kg", "yield_se", "yrs_in_trial")

plot(yield_kg ~ yrs_in_trial, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Year since release",
     ylab = expression("Grain yield ( kg ha"^-1~")"),
     ylim = c(8.0, 12.5))
with(plot_data, segments(x0 = yrs_in_trial, y0=yield_kg + yield_se,
                         y1 = yield_kg - yield_se))
abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000 ,
       b = yrs_coef/1000)
polygon(x = c(yrs_in, rev(yrs_in)),
        y = c(ints[1,],rev(ints[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))
text(5, 8.0, bquote(.(yrs_coef)~kg~ha^-1~year^-1))
mtext("Adjusting for yield erosion", adj = 0)

@
\caption{The relationship between yield and the years since release
  for medium grain rice cultivars. Negative values on the X axis
  indicate experimental yields prior to the official release. The
  estimated erosion of yield over time (averaged over differences in
  cultivar and year) is estimated to be \Sexpr{yrs_coef} kg ha\tsups{-1}
  year\tsups{-1}. Points ($\bullet$) and bars (|) represent the mean and
  standard error of the raw data, respectively. The shaded area is the
  90\% credible interval.}
\label{fig:RES_year}
\end{figure}

There was no evidence of consistent differences between cultivars in
the rate of yield erosion (Model 3), hence the model excluding varying
rates of erosion by cultivar (Model 2) is presented here. The yields
of each cultivar over time were evaluated relative to when the
cultivar was released using the full model which accounts for both
yield erosion and yield improvements over time. The average decrease
in yield over time (i.e., yield erosion) once a cultivar was selected
was estimated at \Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} (90\%
CI: \Sexpr{RES_CI[2]} to \Sexpr{RES_CI[1]}) (Fig \ref{fig:RES_year};
Table \ref{tbl:coef}). The oldest cultivar in the trials
(\Sexpr{gsub("M","M-",unique(mod_data[which.min(mod_data[,"release_yr"]),"id"]))})
is estimated to have lost an average of
\Sexpr{round(yrs_coef*-36/1000,1)} Mg ha\tsups{-1} since its
release. Likewise, an estimated 560 kg ha\tsups{-1} yield
advantage of M-206 over M-202 (the current and former predominant
cultivars, respectively), can be attributed solely to the erosion
of M-202's yield over time (Figs \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}). Further supporting this estimate of yield
erosion, both cultivars yielded similarly at their respective year of
release (Fig. \ref{fig:compare_yrs}), despite the fact that M-202 was
released 19 years prior to M-206 (Fig. \ref{fig:compare_yrs}; Table
\ref{tbl:ids}).

\begin{figure}[h]

<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=
tars <- c("M202","M206")

plot_data = lapply(tars, function(x) {
    dd = mod_data[mod_data$id == x & mod_data$site == "RES",]
    tmp = prep_plot_data(dd$yield_kg / 1000, dd$yrs_in_trial)
    tmp$yrs_in_trial = as.numeric(row.names(tmp))
    tmp
    })

plot(mean ~ yrs_in_trial,
     data = plot_data[[1]],
     xlim = c(-6, 32), ylim = c(6.5,14.5),
     pch = 16, #main = gsub("M","M-", tars[i]),
     ylab = expression("Yield ( kg ha"^-1~")"),
     xlab = "Years after release")
segments(x0 = plot_data[[1]]$yrs_in_trial,
         y0 = plot_data[[1]]$mean + plot_data[[1]]$se,
         y1 = plot_data[[1]]$mean - plot_data[[1]]$se)
points(mean ~ yrs_in_trial,
       data = plot_data[[2]],
       pch = 17, col = "grey40")
segments(x0 = plot_data[[2]]$yrs_in_trial,
         y0 = plot_data[[2]]$mean + plot_data[[2]]$se,
         y1 = plot_data[[2]]$mean - plot_data[[2]]$se,
         col = "grey40")
abline(lm(yield_kg/1000 ~ yrs_in_trial,
          data = mod_data[(mod_data$id %in% tars & mod_data$site == "RES"),]),
       lty = 2, col = "grey50")
legend("topright", legend = gsub("M","M-", tars),
       pch = c(16, 17), col = c("black","grey40"))
text(5, 6.7, bquote(.(yrs_coef)~kg~ha^-1~year^-1))

@
\caption{The performance of two popular cultivars over time. The
  dashed line is the estimated yield erosion as estimated by the full
  model. Points ($\bullet$) represent the mean, and bars (|) the standard
  error of the observed data.}
\label{fig:compare_yrs}
\end{figure}

\subsection{Yield gains}

When yield erosion is not accounted for in the analysis (Model 2),
there is estimated to be an increase per year of approximately
\Sexpr{only_release_coef} kg ha\tsups{-1} year\tsups{-1} (90\% CI
\Sexpr{only_release_CI[1]} to \Sexpr{only_release_CI[2]})
(Fig. \ref{fig:wo_erosion}). However, the full model (Model 1) which
accounts for both the year of release and yield erosion indicates that
the estimated change in grain yield per year in the CCRRF's breeding
trials at the RES was \Sexpr{release_yr_coef} kg ha\tsups{-1}
year\tsups{-1}, with both small positive and negative trends within
the credible interval (90\% CI: \Sexpr{rel_CI[1]} to
\Sexpr{rel_CI[2]}) (Fig. \ref{fig:first_yrs}). According to model
comparison metrics (PS-LOO, WAIC), the full model (Model 1) is
estimated to have superior predictive accuracy compared to the model
which does not account for yield erosion (Model 2)

After correcting for yield erosion, some cultivars were higher
yielding compared to the overall average (as estimated via the random
intercept term in the model). The marginal estimate for the
performance of a cultivar averaged over the uncertainty in the other
parameters in the model (i.e., the random effect) shows three
cultivars with higher than average yields:
\Sexpr{paste(gsub('M','M-',rev(row.names(id_est))[c(1,2,4)]),colla=', ')}
(Figs. \ref{fig:first_yrs} \& \ref{fig:ids_comp}).

\begin{figure}[h]

<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=
yrs <- 1975:2025

pp <- posterior_linpred(RES_rel_fit, re.form = NA,
                        newdata = data.frame(release_yr_c = yrs - 2000,
                                             yrs_in_trial = 0),
                        prob = 0.5)
pp <-  (pp * scl + center) / 1000
ints1 = apply(pp, 2, quantile, p = c(0.05, 0.95))

res = mod_data$site == "RES"
comp_yrs = 2011:2016
idx = mod_data$year %in% comp_yrs

plot_data = prep_plot_data(mod_data$yield_kg[idx] / 1000 ,
                           mod_data$id[idx])

plot_data$tmp = tapply(mod_data$release_yr[idx],
                       mod_data$id[idx], unique)

colnames(plot_data) = c("yield_kg", "yield_se", "release_yr")

# get corrected yield
wo_erosion = expand.grid(id = row.names(plot_data),
            year = comp_yrs,
            release_yr_c = 0,
            yrs_in_trial = 0)
wo_erosion$corr_yield = colMeans(posterior_predict(RES_full,
                               newdata = wo_erosion)) + center/1000
plot_data2 = prep_plot_data(wo_erosion$corr_yield,
                            wo_erosion$id)
plot_data2$release_yr = plot_data$release_yr

plot(yield_kg ~ release_yr, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Release year",
     ylab = "", 
     ylim = c(8, 12.5),
     xaxt = "n",
     xlim = c(1980,2020))
axis(1, at = seq(1980, 2020, by = 10))
mtext(text = expression("Grain yield ( Mg ha"^-1~")"), side = 2, line = 2.5)
abline(a = ((fixef(RES_rel_fit)[1] * scl) + center) / 1000 -
      (2000 * fixef(RES_rel_fit)[2]),
       b = fixef(RES_rel_fit)[2],
       lty = 2)
polygon(x = c(yrs, rev(yrs)),
        y = c(ints1[1,],rev(ints1[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

with(plot_data, segments(x0 = release_yr, y0=yield_kg + yield_se,
                         y1 = yield_kg - yield_se))
text(x = plot_data$release_yr-0.4, y = plot_data$yield_kg + 0.07,
     labels = gsub("M", "M-", row.names(plot_data)),
     pos = 4, cex = 0.75)
text(x = 2000, y = 8, bquote(.(round(fixef(RES_rel_fit)[2] * scl, 1))~kg~ha^-1~yr^-1))
mtext("Without adjusting for yield erosion", adj = 0)

@
\caption{The relationship between yield and the release year of medium
grain rice cultivars for the last 6 years
(\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated without
correcting for yield erosion over time.}
\label{fig:wo_erosion}
\end{figure}

\begin{comment}
\begin{figure}[H]
<<fig=TRUE, echo=FALSE>>=

plot(mean ~ release_yr, data = plot_data2, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Release year",
     ylab = "", #expression("Grain yield -corrected( Mg ha"^-1~")"),
     yaxt = "n",
     ylim = c(8, 12.5),
     xaxt = "n",
     xlim = c(1980,2020))
axis(1, at = seq(1980, 2020, by = 10))
abline(a = ((((fixef(RES_full)[1] * scl) + center) ) -
           (2000 * release_yr_coef)) / 1000,
       b = release_yr_coef/1000,
       lty = 2)
with(plot_data2, segments(x0 = release_yr, y0 = mean + se,
                         y1 = mean - se))
text(x = plot_data2$release_yr-0.4, y = plot_data2$mean + 0.07,
     labels = gsub("M", "M-", row.names(plot_data2)),
     pos = 4, cex = 0.75)
text(x = 2000, y = 8, bquote(.(release_yr_coef)~kg~ha^-1~yr^-1))
mtext("Adjusting for yield erosion", adj = 0)

polygon(x = c(yrs, rev(yrs)),
        y = c(ints2[1,],rev(ints2[2,])),
        border = NA,
        col = rgb(0,0,0,0.1))

@
\caption{The relationship between yield and the release year of medium
  grain rice cultivars for the last 6 years
  (\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated by 
  correcting for yield erosion over time. Yield erosion was corrected
  for by adjusting each observed yield by estimated amount lost 
  (\Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} after release).}
\label{fig:w_erosion}
\end{figure}
\end{comment}

\section{Discussion}

In CA, the cultivars with broad adaptation and adoption over the 33
year period of this study are M-202, M-206, and M-105. Cold tolerance
is a major problem for much of the California rice producing region
~\citep{espe2017}, although it is a bigger problem in the southern
portion of the Sacramento Valley due to the influence of the ocean
winds through the Delta ~\citep{espe16a}. These three cultivars have a
higher degree of cold tolerance, high yield potential, and relatively
short duration and are thus broadly grown with over 70\% of the area
in these cultivars at any given time during the study
period~\citep{ucce}. There are several cultivars that were found to
have better than average yields, even after correcting for yield
erosion. However, these cultivars (M-201, M-204, M-205, and possibly
M-209; Fig. \ref{fig:ids_comp})) are not widely adopted. Importantly,
not all medium grain cultivars that were released were purported to
have higher yields than previous cultivars. Some cultivars had special
traits, but upon release had similar yields to more broadly adapted
cultivars. Examples of this include M-208, which had resistance to
blast disease (\textit{Magnaporthe grisea}) and M-402 which is a
premium quality cultivar. Other cultivars such as M-205 and M-209,
while high yielding, do not have the cold tolerance of the other
cultivars and are thus only recommended for the northern portion of
the Sacramento Valley.

Not accounting for yield erosion, the estimated yield improvement of
\Sexpr{only_release_coef} kg ha\tsups{-1} year\tsups{-1}
(Fig. \ref{fig:wo_erosion}) is lower than the 37 to 60 kg ha\tsups{-1}
year\tsups{-1} estimated by~\citet{Omar2014} in side-by-side
comparisons. However, when accounting for yield erosion and looking
across all cultivars, there is no conclusive evidence of increased
genetic yield potential in medium grain rice cultivars released by
CCRRF's breeding improvement program over the period
\Sexpr{min(mod_data[,'year'])} to \Sexpr{max(mod_data[,'year'])}. 
Grain yield over this period has stayed relatively constant at an average
\Sexpr{round(mean(mod_data[,"yield_kg"])/1000, 1)} Mg
ha\tsups{-1}. Since the estimated yield improvement spans from
\Sexpr{rel_CI[1]} to \Sexpr{rel_CI[2]} kg ha\tsups{-1} year\tsups{-1}
(Fig. \ref{fig:RES_year}; Table \ref{tbl:coef}), both increases and
decreases in the genetic yield potential over the study period are
plausible. This analysis suggests that the CCRRF breeding program has
maintained but not substantially improved the yield
potential in the face of changing biotic and abiotic pressures. This
conclusion holds true even when evaluating the cultivars which have
been breed for broad adaption and high yield potential (M-202, M-206
and M-105). Yields of these cultivars at time of release have all
yielded between
\Sexpr{tmp=mod_data[mod_data[,"yrs_in_trial"]<=0&mod_data[,"id"]%in%c("M202","M206","M105"),];round(min(by(tmp,tmp[,"id"],function(x)mean(x[,"yield_kg"]))/1000),2)}
and
\Sexpr{round(max(by(tmp,tmp[,"id"],function(x)mean(x[,"yield_kg"]))/1000),2)}
Mg ha\tsups{-1} with no indication of yields increasing
(Fig. \ref{fig:first_yrs}).

Yields of a particular cultivar may decline overtime for a number of
reasons. First, ~\citet{Nie2009} reported yield declines in rice due
to continuously flooded conditions over a long time period which could
be overcome with improved N management. This is likely not the case
here as the fields used in this study grow rice only one season a year
and tend to be non-flooded for the remainder of the year. Furthermore,
at this location and unlike most CA rice fields, the fields are left
fallow every other year. Second, changing climate may influence yields
with either warmer~\citep{van2013} or cooler~\citep{espe2017}
temperatures reducing yields. Again, this is unlikely to be the case
in this situation as ~\citet{espe2017} found no significant changes in
growing season temperatures (either high, low, or average temps) the
Sacramento Valley during the study period. Thirdly, small changes in
biotic and abiotic forces have previously been speculated as the cause
for yield erosion of the cultivar IR8 at IRRI~\citep{Peng1998,
Peng2000, Peng2010}. Since the cultivar's genetics remain stable from
the time of selection but the conditions the cultivar experiences over
time are continuously changing, accumulated changes too small to
reliably measure could plausibly lead to yield
erosion~\citep{Peng1998, Peng2000, Peng2010}. However, it is not clear
if this mechanism is responsible for the yield erosion of
\Sexpr{yrs_coef} Mg ha\tsups{-1} year\tsups{-1} found in this
analysis. Biotic stressors such as pests and diseases are low in this
Mediterranean climate. Furthermore, these fields are fallowed every
other year and the residues are burned, further lowering the potential
for pest and disease pressure.

It is striking that the estimated yield erosion is similar to that
reported by \citet{Peng2010} over a similar time span (15\% decline
over roughly 30 years). This suggests the mechanisms responsible could
potentially be endemic to all rice production systems. While efforts
should be made to identify the cause of erosion, this study highlights
both the need for maintenance of plant breeding and the difficulty of
increasing genetic yield ~\citep{espe2017} potential, especially in
systems such as CA where yield is approaching the physiological limit
\citep{espe16a}.

A goal of many breeding programs is to not only maintain the benchmark
yield of the previous ``standard'' but to surpass it, especially if we
are to meet the target of 50\% increased food production by 2050
\citep{Foley2011a}. Given the estimated yield erosion, to hit an
ambitious target of 500 kg ha\tsups{-1} (approximately 5\%) increase
in yield per release (with roughly 4 years between releases), a new
cultivar needs to yield an additional \Sexpr{round(yrs_coef*-4,-1)} kg
ha\tsups{-1} in side-by-side comparison in addition to the 500 kg
ha\tsups{-1} target improvement. Often side-by-side yield improvements
show an improvement in grain yield of 2 to 3\% per release, in which case
no absolute improvements in yield have been made beyond maintaining
yield.

\begin{figure}[h]
<<fig = TRUE, echo = FALSE, width=5, height=5>>=

plot(1, type = "n",
     # xaxt = "n",
     xlab = "Years",
     ylab = expression("Grain yield ( Mg ha"^-1~")"),
     ylim = c(8.5,11.5),
     xlim = c(0,40),
     xaxs = "i")

# arrows(12, 8.2, 27, 8.2, xpd = TRUE, lwd = 2)
abline(a = 10, b = fixef(RES_full)[3], lty = 2, lwd = 2)
text(25, 10.85, "(3) target above maintenance", cex = 0.8,
     srt = 18)#((atan(fixef(RES_full)[3] + 0.06) * 180))/pi)
abline(h = 10, lty = 3, lwd = 2)
text(25, 10.1, "(1) current situation", cex = 0.8)
abline(a = 10, b = fixef(RES_full)[3] + 0.06, lty = 4, lwd = 2)
text(25, 9.35, "(2) no maintenance breeding", cex = 0.8,
     srt = -18)

@ 
\caption{A conceptual diagram highlighting three scenarios: (1)
maintenance breeding, where losses over time are offset by
improvements via plant improvement, (2) no maintenance breeding, where
grain yield slowly erodes over time (as estimated from this study),
and (3) absolute improvement in grain yield (based on ``typical''
yield goals of 3\% increase per release), where gains in grain yield
exceed the losses over time.}
\label{fig:theory}
\end{figure}

Our investigation here is narrowly focused on the rough-rice/field
grain yield. Other advancements due to plant improvement, including
increased quality~\citep{mckenzie1993} and milling
yield~\citep{mckenzie1994}, have been made over the study period and
are important. Additionally, crop duration has been reduced in CA rice
cultivars. For example, while the predominate cultivars M-202 and
M-206 had similar yield potentials at the time of release (Fig. 5),
M-206 matures about 6 days earlier than M-202 (Table
\ref{tbl:ids}). Shorter duration is very important in CA due to an
already short growing season and water limitations (shorter duration
cultivars would presumably require less irrigation water).

California state average yields have increased slowly during the study
period (Fig. \ref{fig:state_yield}) \citep{espe16a}. What are the
potential causes?  It is not due to the adoption of the broadly
adapted cultivars (M-202, M-206, and M-105) as there is no evidence of
increased yield potential among these (Fig. \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}). Although M-205 is not broadly adapted, it has
increase yield potential and its adoption (although limited) in the
northern and warmer part of CA may contribute to some yield gains
reported at the state level. Other factors include improvement in
agronomic management, including more efficient nutrient
management~\citep{Lundy2012, Lundy2015, linquist2009}, and weed
management~\citep{Saito2008, Pittelkow2012, linquist2008}. However,
recent rates of state-wide yield gains are far less than those of the
preceding era, consistent with a system near the yield
ceiling~\citep{Grassini2013}.

It should be noted that while our investigation has been concerned
about increasing yield potential, the estimates here might not capture
yield stability (how consistent grain yields are from year to year). The
occurrence of stress events (e.g., extreme temperatures) are
unpredictable across years and are not well captured by the analysis
here. Yield stability is notoriously difficult to
quantify~\citep[e.g.,][]{lobell2011}, hence further investigation is
needed to explore how yield stability has changed over the study
period.

Planning to meet future demand necessitates increases in grain
yield~\citep{Godfray2010, Foley2011a, Mueller2012}. Meeting the
challenge to produce more grain without expansion of land under
cultivation will rely on new genetics with increased yield
potential. The evidence here suggests that to accomplish true
increases over time, the improvement target needs take into account
erosion in the current ``standard'' cultivars' yield over time
(Fig. \ref{fig:theory}). While side-by-side comparisons will continue
to be useful, comparisons across years to each cultivar's benchmark
yield at the time of release should provide information on the
absolute improvements being made in yield potential over time. New
technologies may need to be adapted to further increase yield
potential, including hybrid rice.

\section{Conclusions}

Here we show that side-by-side cultivar comparisons do not necessarily
indicate gains in yield potential as they do not account for yield
erosion. Quantifying changes in yield potential requires also
examining historical benchmarks. Improvement in grain yield is
essential to meeting the challenges of feeding a growing population in
a changing world. To achieve these improvements, we discuss here that
(1) continuous maintenance via plant improvement is required, without
which yields would decrease over time, (2) the challenge of improving
genetic yield potential is substantial, since absolute gains require
improvement in yield in excess of the maintenance threshold, and (3)
increasing yield through improved management (agronomy) cannot exceed
the ceiling of genetic yield potential. Taken together, these
conclusions underscore both the necessity and substantial challenges
facing rice breeding. Maintaining high quality is imperative for
a breeding program and it is not uncommon in a breeding program
for higher yielding cultivars to be eliminated due to concerns about
quality. However, the goal of increasing yield may require the yield
performance of cultivars to be assessed earlier in the selection
process to avoid discarding these higher yielding entries without full
consideration. We suggest here that historical benchmarks be
used in to evaluate promising rice cultivars in addition to
side-by-side comparisons to maintain perspective of the absolute yield
gains being made.

\newpage

\section{Tables}

<<echo=FALSE, results=tex>>=

CIs = posterior_interval(RES_full)[1:3,] * scl
CIs[1,] = round((CIs[1,] + center)/1000, 1)
CIs = round(CIs, 1)
med_eff = fixef(RES_full) * scl
med_eff[1] = round((med_eff[1] + center)/1000, 1)
med_eff = round(med_eff, 1)

RES_coef = cbind(med_eff, paste(CIs[,1],CIs[,2], sep = " to "))
rownames(RES_coef) = c("Intercept (Mg ha\\tsups{-1})",
                       "Yield improvement (kg ha\\tsups{-1} year\\tsups{-1})",
                       "Yield erosion (kg ha\\tsups{-1} year\\tsups{-1})")
colnames(RES_coef) = c("Median", "90\\% Credible interval")

print(xtable(RES_coef,
             caption = "The estimated median fixed effects and 90\\% credible intervals for the improvement in yield of medium grain rice cultivars developed by the California Cooperative Rice Research Foundation, California, USA over the time period 1984 -- 2016.",
             label = "tbl:coef",
             align = "lcc"),
      caption.placement="top",
      booktabs=TRUE,
      sanitize.colnames.function = function(x){x},
      sanitize.rownames.function = function(x){x})
@
\newpage

<<echo=FALSE, results=tex>>=
ids = aggregate(release_yr ~ id, data = mod_data, unique)
ids$dth = as.integer(round(aggregate(dth ~ id, data = mod_data, mean, na.rm = TRUE)$dth, 0))
colnames(ids) = c("Cultivar", "Release Year", "Days to 50% Heading")
ids = ids[order(ids[,2]),]
ids[,2] = as.integer(ids[,2])
ids[,1] = gsub("M", "M-", ids[,1])

print(xtable(ids,
             caption = "All named medium grain rice cultivars and their associated release year (from 1984 -- 2015) planted at the Rice Experiment Station (RES), Biggs, California.",
             label = "tbl:ids"),
      include.rownames = FALSE,
      booktabs = TRUE,
      caption.placement = "top")

@

% Notes: Duration has gotten shorter. Etc. Rearrange - End
% introduction with evidence of erosion, quatify the effect on yield
% increases.

\newpage

\section{References}\label{references}

\setlength{\parindent}{-0.2in} \setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}

\bibliography{breeding}

\newpage

\section{Supplemental tables and
figures}\label{supplemental-tables-and-figures}

\beginsupplement
\begin{comment}
\begin{figure}[H]
<<fig=TRUE, width = 8, height = 5, out.width = '85%', echo=FALSE>>=
tars <- c("M202","M206")
par(mfrow = c(1, length(tars)),
    mar = c(4,0,2,0),
    oma = c(4,4,0.5,0.5))

plot_data = lapply(tars, function(x) {
    dd = mod_data[mod_data$id == x & mod_data$site == "RES",]
    tmp = prep_plot_data(dd$yield_kg/1000 , dd$year)
    tmp$year = as.numeric(row.names(tmp))
    tmp
    })

invisible(sapply(1:2, function(i){
    plot(mean ~ year,
         data = plot_data[[i]],
         xlim = c(1985, 2016), ylim = c(7.5,13.5),
         pch = 16, main = gsub("M","M-", tars[i]),
         ylab = "",
         yaxt = "n",
         xlab = "Year",
         cex.axis = 0.8)
    segments(x0 = plot_data[[i]]$year,
             y0 = plot_data[[i]]$mean + plot_data[[i]]$se,
             y1 = plot_data[[i]]$mean - plot_data[[i]]$se)
    if(i == 1){
        axis(2, cex.axis = 0.8,)
        mtext(expression("Yield ( Mg ha"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }
    # abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000 -
    #            (2000 * yrs_coef/1000),
    #        b = yrs_coef/1000,
    #        lty = 1)
    lines(as.numeric(rownames(plot_data[[i]])),
          11.2 + (1:nrow(plot_data[[i]])) * (yrs_coef/1000))
    # abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000,
           # b = yrs_coef/1000)

    # abline(h = mean(mod_data$yield_kg[(mod_data$id == tars[i] &
    #                                    mod_data$site == "RES")] ),
    #        lty = 1, col = "grey50")
    # mod = lm(mean ~ year, data = plot_data[[i]])
    # tmp = predict(mod)
    # if(i == 1){
    #     abline(mod)
    # }else{
    #     lines(as.numeric(names(tmp)), tmp)
    # }
    abline(h = 11.2, lty = 2)
    # text(2000, 7.5, bquote(.(round(coef(mod)[2] * 1000, 1))~kg~ha^-1~year^-1))
}))

@
\caption{The performance of two major cultivars over the period
\Sexpr{min(mod_data[,'year'])} -- \Sexpr{max(mod_data[,'year'])}. The
solid line is the estimated yield erosion (\Sexpr{yrs_coef} kg
ha\tsups{-1} year\tsups{-1}), while the dashed line is the published
3-year average yield of M-202 at its release (1982-1984). Both
cultivars have a similar yield at their respective year of release,
and similar decreases over time.}
\label{fig:compare_yrs2}
\end{figure}
\newpage
\end{comment}

% \begin{figure}[H]
% <<fig=TRUE, echo=FALSE>>=
% tars <- c("M202","M206")
% comp_yrs = 2004:2005

% boxplot(yield_kg/1000 ~ id, data = mod_data[(mod_data$id %in% tars) &
%                                             (mod_data$year %in% comp_yrs),],
%         ylab = expression("Grain yield ( Mg ha"^-1~")"))

% @
% \caption{An example of a typical side-by-side comparison between a older (M-202)
%   and newer (M-206) cultivar. Shown is the performance of these cultivars at the CA Rice
%   Experiment Station (RES) over the period \Sexpr{min(comp_yrs)} --
%   \Sexpr{max(comp_yrs)}. The newer
%   cultivar, M-206, appears to have a yield advantage over the older
%   cultivar, M-202, though this comparison does not account for the 10
%   year difference in years since release between the cultivars.}
% \label{fig:compare_vars}
% \end{figure}
% \newpage

\begin{figure}[H]
<<echo = FALSE, fig = TRUE, width = 7, height = 6>>=
library(ggplot2)
library(ggthemes)
ggplot(mod_data, aes(x = year, y = yield_kg/1000)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_smooth(method = "lm", color = "black") +
    theme_few() +
    xlab("Year") +
    ylab(expression("Grain yield ( Mg ha"^-1~")")) + 
    facet_wrap(~id, ncol = 4)

@
\caption{The yield of all medium grain released cultivars at the
  California Cooperative Rice Research Foundation Rice Experiment
  Station over the period \Sexpr{min(mod_data[,'year'])} --
  \Sexpr{max(mod_data[,'year'])}.}
\label{fig:all_cult}
\end{figure}

\newpage
\begin{figure}[ht]
<<echo=FALSE, fig = TRUE>>=

dotchart(id_est[,1],
         labels = gsub("M", "M-", row.names(id_est)),
         pch = 16,
         xlim = c(min(id_est[,2]),max(id_est[,3])),
         xlab = expression("Estimated difference from average ( Mg ha"^-1~")"))
segments(x0 = id_est[,2], x1 = id_est[,3], y0 = 1:nrow(id_est))
abline(v = 0, lty = 2)
@
\caption{The marginal estimates for the difference between cultivars,
  averaged over the uncertainty in the fixed effect parameters (i.e.,
  yield erosion and annual yield gains).}
\label{fig:ids_comp}
\end{figure}

\newpage

\end{document}